### 2022-04-07 manuscript/header.tex
initial commit

\usepackage{amsmath}

% ---------- CHOOSE A FONT ----------

\usepackage{fontspec}
%\setmainfont{Fira Sans}
\setmainfont{Verdana}

\usepackage[english]{babel}							   % hyphenation etc for English

% --------- MISC FORMATTING ---------
%
\usepackage{setspace}                % needed for doublespacing
%\doublespacing                       % doublespaced line spacing
\usepackage{graphicx}                % for importing graphics into figures
\usepackage{float}                   % better control of figure placement
\usepackage{hyperref}                % for clickable URLs and email addresses
\usepackage{cleveref}
\urlstyle{sf}                        % san-serif font for URLs
\usepackage[margin=1.2in]{geometry}  % control page margins
\usepackage[labelfont=bf]{caption}   % make caption labels boldface
\usepackage[bottom]{footmisc}        % footnotes at bottom of page
\setlength{\skip\footins}{10mm}      % obsessing about footnote spacing
\setlength{\parskip}{1ex}            % space between paragraphs
%\setlength{\parindent}{3em}	     % paragraph indentation
\usepackage{authblk}                 % author and affiliation formatting
\renewcommand\Affilfont{\small}
\usepackage{booktabs}
\usepackage{pdflscape}

% Biblio
\usepackage[style=apa,
labeldate=year,
isbn=false,
eprint=false,
date=year,
url=false,
backend=biber]{biblatex}
\bibliography{references.bib}

%-----------
% Line numbers
\usepackage{lineno}                  % add line numbers to margin
\def\linenumberfont{\normalfont\footnotesize\sffamily} % line numbers
\setlength\linenumbersep{9mm}                          % line numbers
\linenumbers                                           % line numbers

### 2022-04-07 manuscript/ms_genomes.Rmd
initial commit

---
title: "Draft genomes of the three northern hemisphere blue mussel species: *Mytilus edulis*, *Mytilus galloprovincialis* and *Mytilus trossulus*"
author:
  - Alexis Simon^1^*
  - Christine Arbiol^1^
  - Nicolas Bierne^1^
date: '`r Sys.Date()`'
output:
  bookdown::pdf_document2:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    df_print: kable
    keep_tex: yes
    number_sections: yes
    toc: no
    includes:
      in_header: header.tex
  html_document:
    toc: no
documentclass: article
# classoption: twocolumn, landscape, etc
papersize: a4
pagestyle: plain  # plain, empty, headings
geometry: margin=1in
#linestretch: 2
fontfamily: mathpazo
fontsize: 11pt
bibliography: 
  - references.bib
  - knitcitations.bib
# zotero: "My shared Zotero group"
csl: g3.csl
# csl: https://www.zotero.org/styles/ecology
link-citations: yes
linkcolor: RoyalBlue
urlcolor: RoyalBlue
links-as-notes: false
editor_options: 
  markdown: 
    wrap: sentence
---

<!--# Insert author affiliations below -->

\small

^1^ CNRS, Univ Montpellier, ...

`*` Corresponding author: [alexis.simon\@normalesup.org](mailto:alexis.simon@normalesup.org){.email}

\normalsize

<!--# Write Abstract and Keywords below -->

```{=tex}
\vspace{1cm}
\hrule
```
Using the 10X chromium long reads technology, we provide draft genomes for three closely related blue mussel species from the *Mytilus* species complex.
The objective was to produce affordable genomic resources for population and evolutionary genomic studies.
Genomes are fragmented but represent a large portion of the genome, with good sizes and BUSCO scores.

```{=tex}
\vspace{3mm}
\hrule
\vspace{5mm}
```
*Keywords*: *Mytilus edulis*, *Mytilus galloprovincialis*, *Mytilus trossulus*, Genome assembly, 10X chromium

```{=tex}
\bleft
\newpage
```
```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

### Chunk options: see http://yihui.name/knitr/options/ ###
## Text results
opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
# opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "output/figures/")
```

```{r knitcitations, cache = FALSE}
library(knitcitations)
cleanbib()
cite_options(citation_format = "pandoc")
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
```

<!--# Start writing your paper below -->

# Introduction

Write your introduction here.

------------------------------------------------------------------------

# Methods

## Biological material and DNA extraction

One individual for each species of interest was collected and processed fresh.

Collection locations: - *Mytilus galloprovincialis* Sète, France, - *M. edulis* ???
- *M. trossulus* ???

Whole mussels were placed in 50 mL falcon tubes containing 25 ml of TNES-Urea solution and incubated for 4-6 weeks at room temperature (TNES-Urea: 10 mM Tris-HCl pH 7.4; 120 mM NaCl; 10 mM EDTA pH 8.0; 0.5% SDS; 4 M urea).

After this period of pre-treatment at ambient temperature, proteinase K was added at a final concentration of 150 $\mu$g/mL and the solution was incubated overnight at 56°C.

High Molecular weight genomic DNA was extracted following (From Nakayama et al., 1994, Chromosoma).
We used a 15 mL Phase Lock Gel Heavy extraction with 3 steps of phenol/chloroform/isoamylalcool (25/24/1) Tris pH 8,1 followed by 2 chloroform extractions.
After the last extraction, the aqueous supernatant was precipitated with 2 volumes of 100% EtOH and the pellet was hooked from the solution with a sterile glass Pasteur pipette.
The pellet was rinsed several times in 80% EtOH before being dried at room temperature.

DNA was resuspended with an appropriate volume of biomolecular water at 65°C for several hours, the duration of this incubation depending on the size of the granules.
DNA was then stored at 4°C.

Prior to the construction of the DNA libraries, DNA was repaired with NEBNext FFPE DNA Repair Mix according to the manufacturer's instructions.

## Library preparation and sequencing

The 10X chromium library preparation and sequencing was subcontracted to the MGX platform (Montpellier, France).

They followed 10X Genomics Genome Reagen Kit (*Genome Solution*) protocol to produce the 10X linked reads library using a Chromium microfluidic chip.

The library was sequenced on an Illumina NovaSeq 6000 with an S4 flowcell.

## Draft genome assemblies

Explain here v1, v4, v5

-   v1: `Supernova` v2.1.1
-   v4: `purge_dups` (`longranger align`)
-   v5: `agouti` using RNAseq data (`augustus` to get gff3, RNAseq `rcorrector` + `trimgalore` + `bwa-mem2` + merge bams). RNAseq data are from both NCBI and Maurine's/Erika's experiments.
-   v6: blobtoolkit filtration based on groups

-\> gives mtro_01, medu_01, mgal_01

## *M. trossulus* assembly improvement using minion reads and *M. coruscus* assembly

-   `LRScaf` using ONT reads (`minimap2` alignment)
-   `RagTag` using *M. coruscus* chromosomal assembly
-   `Pilon` using PE reads from 10X library (debarcoded + filtered with `fastp`, `bwa-mem2` alignement) + ONT reads

-\> gives mtro_02 (tros_v7)

## Repeats ???

## Annotation ???

## Document softwares, versions and DOIs:

-   Supernova v2.1.1 [cite](https://doi.org/10.1101/gr.214874.116)
-   Longranger v2.2.2
-   proc10xG <https://github.com/ucdavis-bioinformatics/proc10xG> commit 7afbfcf
-   nubeam-dedup <https://github.com/daihang16/nubeamdedup> commit 25dd385 [cite](https://doi.org/10.1093/bioinformatics/btaa112)
-   KMC <https://github.com/tbenavi1/KMC> commit 1df71f6
-   GenomeScope 2.0 <https://github.com/tbenavi1/genomescope2.0> commit 5034ed4 [cite](https://doi.org/10.1038/s41467-020-14998-3)
-   (? Smudgeplot commit 950d7b5)
-   Purge_Dups <https://github.com/dfguan/purge_dups> commit e1934bb [cite](https://doi.org/10.1093/bioinformatics/btaa025)
-   AGOUTI <https://github.com/svm-zhang/AGOUTI> commit a7e65d6 [cite](https://doi.org/10.1186/s13742-016-0136-3)
-   augustus v3.3.3 (for AGOUTI analysis) [@Stanke2008]
-   Blobtoolkit v2.4.0 [cite](https://doi.org/10.1534/g3.119.400908)
-   KAT v2.4.2 [cite](https://doi.org/10.1093/bioinformatics/btw663)
-   fastp v0.20.1 [@Chen2018a]
-   samtools v1.12 [@Li2009]
-   minimap2 v2.17 [@LiLi2017]
-   bwa-mem2 v2.2.1 [@VasimuddinMd2019]
-   bedtools v2.29.2 [@Quinlan2010]
-   seqkit v0.13.2 [@Shen2016]
-   rcorrector v1.0.4
-   trim-galore v0.6.6 [cite](https://github.com/FelixKrueger/TrimGalore) (Cutadapt + FastQC wrapper)
-   busco v5.1.1 [@Seppey2019]
-   assembly_stats v0.1.4 [cite](https://doi.org/10.5281/zenodo.3968774)
-   ragtag v1.1.1 [@Alonge2019]
-   pilon v1.24 [@Walker2014]
-   TETools DFAM container v1.3.1
-   RepeatModeler v2.0.1 [@Flynn2020]
-   cd-hit v4.8.1 [@Fu2012]
-   RepeatMasker v4.1.2-p1 (Smit, AFA, Hubley, R. RepeatModeler Open-1.0. 2008-2015 <http://www.repeatmasker.org>) [@Smit2013]

------------------------------------------------------------------------

# Results and discussion

Trees in forest *A* grew taller than those in forest *B* (mean height: `r mean(25, 31, 28)` versus `r mean(13, 19, 16)` m).

And many more cool results that get updated dynamically, e.g. see Table \@ref(tab:Table-mtcars) and Fig.
\@ref(fig:scatterplot).
Note Tables and Figures are **cross-linked and numbered automatically**.
They could also appear in the middle of the document, not necessarily at the end.

See also Fig.
\@ref(fig:SupplFig) and Table \@ref(tab:SupplTable) in the Supplementary Material.

```{r}
# asm_show <- c(
#   "coruscus_GCA017311375", 
#   "gallo_GCA900618805",
#   "edu_GCA905397895",
#   "eduam_GCA019925415",
#   "gallo_v6",
#   "gallo_v7",
#   "edu_v6", 
#   "edu_v7",
#   "tros_v6", 
#   "tros_v7"
# )
```

## Assembly statistics

```{r, include=FALSE}
source("../scripts/asm_stats_prep.R", local = knitr::knit_global())
```

```{r table_asm_stats}
asm_stats_table %>%
  kable(.)
```

## BUSCO scores

```{r load_busco}
busco <- read_tsv("data/busco_summary.tsv", col_types = cols()) %>% 
  mutate(asm = str_replace(asm, "_metazoa_odb10|_mollusca_odb10", "")) %>% # clean asm names
  mutate(cat = fct_relevel(cat, "M", "F", "CD", "CS")) %>%
  mutate(asm = fct_relevel(asm, asm_show) %>% 
           fct_rev()) %>% 
  group_by(asm, db) %>% 
  mutate(proportion = N/cur_data()$N[cur_data()$cat=="T"]) %>% 
  mutate(pct_fmt = paste(format(round(proportion*100, digits = 2)), "%"))
```

```{r full_figure_mollusca, fig.width=10, fig.height=4}
busco %>% 
  filter(db == "mollusca_odb10") %>% 
  filter(cat %in% c("CS", "CD", "F", "M")) %>%
  filter(asm %in% asm_show) %>% 
  ggplot(., aes(x = N, y = asm, fill = cat)) + 
    geom_bar(position = "fill", stat = "identity", width = 0.9) +
    scale_fill_manual(values = c("#D55E00", "#E69F00", "#56B4E9", "#009E73"),
                      guide = guide_legend(reverse=T),
                      name = "categories", 
                      labels = c("Missing", "Fragmented", "Complete duplicated", "Complete single copy")) +
    scale_x_continuous(expand=c(0,0), labels = scales::percent, breaks = c(0,0.2,0.4,0.6,0.8,1)) + 
    scale_y_discrete(expand=c(0,0)) +
    geom_text(aes(label = N), position = position_fill(vjust = .5), size = 3) +
    labs(x = "BUSCO percentage", y = NULL, title = "BUSCO scores Mollusca_odb10 (5295 genes)") +
    theme_minimal() +
    theme(axis.ticks.x = element_line(),
          panel.grid = element_blank(),
          legend.position = "bottom",
          plot.margin = margin(t=5,r=15,b=0,l=5))
```

```{r full_figure_metazoa, fig.width=10, fig.height=4}
busco %>% 
  filter(db == "metazoa_odb10") %>% 
  filter(cat %in% c("CS", "CD", "F", "M")) %>%
  filter(asm %in% asm_show) %>% 
  ggplot(., aes(x = N, y = asm, fill = cat)) + 
    geom_bar(position = "fill", stat = "identity") +
    scale_fill_manual(values = c("#D55E00", "#E69F00", "#56B4E9", "#009E73"),
                      guide = guide_legend(reverse=T),
                      name = "categories", 
                      labels = c("Missing", "Fragmented", "Complete duplicated", "Complete single copy")) +
    scale_x_continuous(expand=c(0,0), labels = scales::percent, breaks = c(0,0.2,0.4,0.6,0.8,1)) + 
    scale_y_discrete(expand=c(0,0)) +
    geom_text(aes(label = N), position = position_fill(vjust = .5), size = 3) +
    labs(x = "BUSCO percentage", y = NULL, title = "BUSCO scores Metazoa_odb10 (954 genes)") +
    theme_minimal() +
    theme(axis.ticks.x = element_line(),
          panel.grid = element_blank(),
          legend.position = "bottom",
          plot.margin = margin(t=5,r=15,b=0,l=5))
```

# Conclusion

Wrap up

# Acknowledgement

On the shoulders of giants.

# References

```{r write_citations, cache=FALSE, include=FALSE}
#write.bibtex(file = "knitcitations.bib")
```

::: {#refs}
:::

\eleft

\clearpage

<!--# Tables below -->

\listoftables

\newpage

```{r Table-Iris}
kable(head(iris), caption = "A glimpse of the famous Iris dataset.", booktabs = TRUE)
```

\clearpage

```{r Table-mtcars}
kable(mtcars[10:16, ], caption = "Now a subset of mtcars dataset.", booktabs = TRUE)
```

\clearpage

<!--# Figures below -->

\listoffigures

\newpage

```{r scatterplot, fig.cap="Just my first figure with a very fantastic caption."}
x <- rnorm(100)
y <- jitter(x, 1000)
plot(x, y)
```

\newpage

\blandscape

```{r Fig-landscape, fig.cap="Second figure in landscape format."}
a <- sort(rnorm(100))
b <- c(rep("Group Small", 35), rep("Group Big", 65))
boxplot(a ~ b)
```

\elandscape

\clearpage

# SUPPLEMENTARY MATERIAL {.unnumbered}

\beginsupplement

# Other things we tried

## Raw reads filtration

Explain pipeline, it did not improve assembly -\> v2, v3

\clearpage

```{r SupplTable}
kable(head(iris), caption = "A supplementary table.", booktabs = TRUE)
```

\clearpage

```{r SupplFig, fig.cap="A Supplementary Figure."}
x <- rnorm(100)
y <- jitter(x, 1000)
plot(x, y)
```

\newpage

```{r sessioninfo, eval = FALSE}
# set eval = FALSE if you don't want this info (useful for reproducibility) to appear 
sessionInfo()
# sessioninfo::session_info()
```

### 2022-04-07 manuscript/ms_genomes.tex
initial commit

% !TeX encoding = UTF-8
% !TeX spellcheck = en_US
% !TeX program = xelatex
\documentclass[11pt, a4paper]{article}

\include{header.tex}

\title{Draft genomes of the three northern hemisphere blue mussel lineages: North and South European \emph{Mytilus edulis} and Mediterranean \emph{Mytilus galloprovincialis}.}

\author[1,*]{Alexis Simon}
\author[1]{Christine Arbiol}
\author[1]{Nicolas Bierne}

\affil[1]{CNRS, Univ Montpellier, \ldots{}}

%

\begin{document}

\maketitle

\vspace{1cm}
\hrule
\begin{abstract}
	Using the 10X chromium long reads technology, we provide draft genomes for three closely related blue mussel species from the \emph{Mytilus} species complex.
	The objective was to produce affordable genomic resources for population and evolutionary genomic studies.
	Genomes are fragmented but represent a large portion of the genome, with good sizes and BUSCO scores.
\end{abstract}

\vspace{3mm}
\hrule
\vspace{5mm}

\noindent
\textbf{Keywords}: \emph{Mytilus edulis}, \emph{Mytilus galloprovincialis}, Genome assembly, 10X chromium

\noindent
$^{*}$ \textbf{Corresponding author}: \href{mailto:alexis.simon@normalesup.org}{\nolinkurl{alexis.simon@normalesup.org}}

\newpage

\section{Rationale and objectives}\label{introduction}

The \textit{Mytilus} species complex has been a model system in population genetics, adaptation, hybridization and speciation since genetic variants could be identified \parencite{Milkman1970, Koehn1972, Ahmad1977, Skibinski1978, Quesada1995b, Bierne2003, Fraisse2016a, Simon2021}.

The \textit{Mytilus} species complex is composed of three taxonomically recognized and partially reproductively isolated species in the northern hemisphere,
\textit{Mytilus edulis}, \textit{M. galloprovincialis} and \textit{M. trossulus}.
Within each species, evolutionary relevant lineages can be identified.
Two lineages of \textit{M. galloprovincialis} can be identified, one Atlantic lineage and one Mediterranean lineage separated by a hybrid zone along the Almeria-Oran front \parencite{Quesada1995, Fraisse2016a, ElAyari2019}.
Three lineages of \textit{M. edulis} can be identified: (i) an American lineage, (ii) a Southern European lineage, and (iii) a Northern European lineage \parencite{Fraisse2016a, Simon2020}.

Therefore, need for diverse references among the complex...

\section{Methods}\label{methods}

\paragraph{General notes:}
The entire assembly was carried out using a Snakemake \parencite{Molder2021} pipeline available on github at \url{https://github.com/alxsimon/assembly_10x}.
Where deemed important, parameters are given in the text.
For brevity and simplicity, not all information might be available in the text.
However, all parameters, software versions and steps are retrievable from the repository.

\paragraph{Important caveat:}
The assembled genome of MeduEUN (\textit{M. edulis} Northern lineage) was initially thought to be \textit{M. trossulus}.
Therefore some assembly and annotation steps wrongly used \textit{M. trossulus} transcriptomes.
While this is not ideal, we think results have not been strongly impacted by this issue.

\subsection{Biological material and DNA extraction}

One individual for each species of interest was collected and processed fresh.

Collection locations:
\begin{itemize}
	\item \emph{M. galloprovincialis} Mediterranean MgalMED; Sète, France
	\item \emph{M. edulis} Southern European MeduEUS; ???
	\item \emph{M. edulis} Northern Europe MeduEUN; ???
\end{itemize}

Whole mussels were placed in 50 mL falcon tubes containing 25 ml of TNES-Urea solution and incubated for 4-6 weeks at room temperature (TNES-Urea: 10 mM Tris-HCl pH 7.4; 120 mM NaCl; 10 mM EDTA pH 8.0; 0.5\% SDS; 4 M urea).

After this period of pre-treatment at ambient temperature, proteinase K was added at a final concentration of 150 \(\mu\)g/mL and the solution was incubated overnight at 56°C.

High Molecular weight genomic DNA was extracted following \textcite{Nakayama1994}.
We used a 15 mL Phase Lock Gel Heavy extraction with 3 steps of phenol/chloroform/isoamylalcool (25/24/1) Tris pH 8,1 followed by 2 chloroform extractions.
After the last extraction, the aqueous supernatant was precipitated with 2 volumes of 100\% EtOH and the pellet was hooked from the solution with a sterile glass Pasteur pipette.
The pellet was rinsed several times in 80\% EtOH before being dried at room temperature.

DNA was resuspended with an appropriate volume of biomolecular water at 65°C for several hours, the duration of this incubation depending on the size of the granules.
DNA was then stored at 4°C.

Prior to the construction of the DNA libraries, DNA was repaired with NEBNext FFPE DNA Repair Mix according to the manufacturer's instructions.

\subsection{Library preparation and sequencing}

The 10X chromium library preparation and sequencing was subcontracted to the MGX platform (Montpellier, France).
The 10X linked reads libraries for each individual were produced following the 10X Genomics Genome Reagen Kit (\emph{Genome Solution}) protocol using a Chromium microfluidic chip.
Libraries were subsequently sequenced on an Illumina NovaSeq 6000 with an S4 flowcell.

\subsection{Preprocessing 10X reads}\label{preproc}

We preprocessed 10X reads using the following pipeline for use in several algorithms (\cref{draft_asm}).
We first removed duplicate reads using \texttt{Nubeam-dedup} (\cite{Dai2020}; commit 25dd385).
We then used \texttt{proc10xG} (\url{https://github.com/ucdavis-bioinformatics/proc10xG}; commit 7afbfcf) to split the reads from their 10X barcodes for further processing.
We used a custom filtering step designed to remove under- and over-represented barcodes and their associated reads from the data (\texttt{filter\_barcodes.R} script, see \cite{Bary\&Gagnaire??}).
[[Add supfigs of filter\_barcodes]].
Reads were additionally filtered with \texttt{fastp} (v0.20.1; \cite{Chen2018}) with the main objective to remove poly-G tails created by the Illumina Novaseq sequencing technology.
We obtained at this point reads equivalent to a short-read sequencing run, usable in some parts of the assembly and quality control pipeline.
Additionally, to obtain reads compatible to 10X genomics tools, we filtered and reassembled reads with their 10X barcodes in \texttt{proc10xG} using the \texttt{filter\_10xReads.py} and \texttt{regen\_10xReads.py} scripts.

\subsection{Initial genome assemblies}\label{draft_asm}

For each genome we used \texttt{Supernova} v2.1.1 \parencite{Weisenfeld2017} to assemble raw 10X reads.
To avoid hard stops in \texttt{Supernova} due to both data quantity slightly under 10X genomics recommendations and an overestimation of genome size by Supernova, we used all available reads (\texttt{--maxreads='all'}) and accepted extreme coverage (\texttt{--accept-extreme-coverage}).
We produced every style of Fasta output available in \texttt{Supernova} but only used the pseudo-haploid output in the following pipeline.

To remove duplicate haplotypes in the assemblies, we followed the \texttt{purge\_dups} pipeline (\cite{Guan2020}; commit e1934bb).
We first used the \texttt{longranger} v2.2.2 align algorithm to map preprocessed reads (\cref{preproc}) to the pseudo-haploid genomes to use in the \texttt{ngscstat} step.
\texttt{Minimap2} (v2.17; \cite{Li2017}) was used in the contig self-mapping step.
We obtained the purged assemblies using the \texttt{get\_seqs} steps without restricting the purging to the end of contigs (without the \texttt{-e} option).

We used the program \texttt{AGOUTI} (\url{https://github.com/svm-zhang/AGOUTI}; commit a7e65d6; \cite{Zhang2016a}) to improve scaffolding by using paired end RNA-seq reads.
For each species, a different set of published transcriptomes were used (Supplementary Table [[...]] for accessions).
\texttt{AGOUTI} require a gene prediction as input in addition to RNA-seq reads.
We used \texttt{Augustus} (v3.3.3; \cite{Stanke2008}) to produce an intermediate annotation for each assembly using the \textit{Caenorhabditis} model.
RNA-seq reads were first cleaned using \texttt{rcorrector} (v1.0.4; \cite{Song2015}) and \texttt{trimgalore} (v0.6.6; \url{https://www.bioinformatics.babraham.ac.uk/projects/trim_galore}; \texttt{--quality 20 --stringency 1 -e 0.1 --length 70}).
RNA-seq reads were mapped independently using \texttt{bwa-mem2} (v2.2.1; \cite{VasimuddinMd2019}), and then merged as a common bam file for each assembly with \texttt{samtools} (v1.12; \cite{Li2009}).
Finally, the \texttt{AGOUTI} scaffolding pipeline was run using the gene prediction and mapped RNA-seq reads (-minMQ 20 -maxFracMM 0.05; using \texttt{python} v2.7.15 and \texttt{samtools} v1.10 for compatibility).

As a last step, we ran \textit{Blobtoolkit} (v2.4.0; \cite{Challis2020}) on the three assemblies to evaluate quality and potential contamination levels.
We used a custom script (\texttt{btk\_conta\_extraction.py}) to filter the assembly contigs based on the taxids found by the \textit{Blobtoolkit} pipeline to remove the most obvious contaminations.
Contigs matching taxids associated with viruses, bacteria and non-mollusca eukaryotes were removed.
More specifically contigs for were removed for virus contamination if they presented a hit percentage of more than 10~\% of their length.
Contigs were removed for eukaryote contamination only when presenting only hits to taxa outside Mollusca on more than 10~\% of their length.
The list of retained contigs was used to filter the fasta assembly file using \texttt{seqkit} (v0.13.2; \cite{Shen2016}).

\subsection{Scaffolding on the \textit{Mytilus coruscus} genome}

At the time of assembly, the closest relative of the \textit{Mytilus} species of interest with a chromosome scale assembly was \textit{Mytilus coruscus} (GCA\_017311375.1; \cite{Yang2021}).
Given the conserved number of 14 chromosomes in the \textit{Mytilus} genus, we decided to scaffold our contigs on this high quality reference.
Additionally, we had Oxford Nanopore reads for the MeduEUN individual. 
We used \texttt{minimap2} (v2.17; \cite{Li2017}) and \texttt{LRScaf} (v1.1.10; \url{https://github.com/shingocat/lrscaf}) to first improve the MeduEUN assembly with this small amount of long reads.

Then for each assembly, we ran the scaffolder \texttt{RagTag} (v1.1.1; \cite{Alonge2021}) to position contigs on the \emph{M. coruscus} chromosomal assembly.

A final polishing step was performed using \texttt{Pilon} (v1.24; \cite{Walker2014}).
\texttt{Pilon} attempts to improve the assembly based on mapped read information (gap filling and error corrections).
We used \texttt{bwa-mem2} to map the debarcoded and filtered 10X reads (\cref{preproc}).
In addition, Oxford Nanopore reads for MeduEUN were also used in \texttt{Pilon} for the given assembly.

\subsection{Repeats}

We masked repeats for the purpose of annotation using \texttt{RepeatModeler} (v2.0.1; \cite{Flynn2020}) and \texttt{RepeatMasker} (v4.1.2-p1; \cite{Smit2013}) through the TETools DFAM container (v1.3.1; \url{https://hub.docker.com/r/dfam/tetools}).
We first built repeat databases for each of five assemblies MgalMED, MeduEUS, MeduEUN, \textit{M. coruscus} (GCA\_017311375.1; \cite{Yang2021}) and \textit{M. galloprovincialis} from the Atlantic (GCA\_900618805.1; \cite{Gerdol2020}).
Then, we built a common \textit{Mytilus} database of repeats by merging those five databases using \texttt{cd-hit} (v4.8.1; \cite{Fu2012}). We used the same options as used by default in \texttt{RepeatModeler}: \texttt{-aS 0.8 -c 0.8 -g 1 -G 0 -A 80 -M 10000}.
Finally, soft repeat masking was performed on the assemblies with \texttt{RepeatMasker} using the merged database.

\subsection{Annotation}

To obtain an annotation for each assembly, we used both database and RNA-seq information.
We used \texttt{Braker2} (v2.1.6; \cite{Bruna2021}) as an initial step, using both a protein database and RNA-seq reads (preprocessed in \cref{draft_asm}).
To build the protein database, we used all \textit{Mollusca} proteins (taxid 6447) from OrthoDB (v10.1; \cite{Kriventseva2019}).
To provide gene presence hints, we mapped all RNA-seq reads for each species using \texttt{HISAT2} (v2.2.1; \cite{Kim2019}).

We used the \texttt{Mantis} pipeline (\url{https://github.com/PedroMTQ/mantis}; commit c6cb597; \cite{Queiros2021}) to obtain consensus annotations of genes based on multiple databases.
Protein sequences for each assembly were built from the \texttt{Braker2} annotation using the \texttt{python} module \texttt{gff3tool} (v2).

\subsection{NCBI submission}

The NCBI submission process identified a few errors that needed correcting.
A small number of adaptor sequences and duplicates were removed to comply with NCBI requirements
(see the rule \texttt{ncbi\_submission\_changes.smk} in the pipeline for more details).
Assemblies are available under the following accessions: 
JAKGDF000000000 for MgalMED,
JAKGDG000000000 for MeduEUS,
and JAKGDH000000000 for MeduEUN.

\subsection{Quality assessments and comparisons}

Preprocessed 10X reads (\cref{preproc}) were used to first estimate estimate genome size and heterozygosities of the three individuals.
We used the reference free k-mer based method \texttt{genomescope} (\url{https://github.com/tbenavi1/genomescope2.0}; commit 5034ed4; \cite{Ranallo-Benavidez2020}) and the fork of the \texttt{KMC} k-mer  counting program (\url{https://github.com/tbenavi1/KMC}; commit 1df71f6).

Assembly statistics were computed with the python module \texttt{assembly\_stats} (v0.1.4; \cite{Trizna2020}).

To assess the remaining level of duplication in the assemblies, we used the program \texttt{KAT} (v2.4.2; \cite{Mapleson2017}) to compare k-mer spectra from reads (preprocessed 10X) and from the assembly.

Finally, gene completion analyses were carried out using \texttt{BUSCO} (v5.1.1; \cite{Manni2021}).
We a Metazoan (\texttt{metazoa\_odb10.2021-02-24}) and a Molluscan database (\texttt{mollusca\_odb10.2020-08-05}) to assess and compare new and published assemblies.
Published assemblies are:
\begin{itemize}
	\item \textit{M. coruscus}, GCA\_017311375.1, \textcite{Yang2021};
	\item \textit{M. galloprovincialis} from the Altantic lineage, GCA\_900618805.1, \textcite{Gerdol2020};
	\item \textit{M. edulis} from the Southern European lineage, GCA\_905397895.1, \textcite{Corrochano-Fraile2021}
	\item \textit{M. edulis} from the American lineage, GCA\_019925275.1.
\end{itemize}

\subsection{Phylogenetic species tree}

We compiled protein sequences from published \textit{Mytilus} genomes and transcriptomes,
in addition to the current three genomes and annotations, to build a species tree.
We used transcriptomes produced in \textcite{Popovic2020a} for American \textit{M. edulis}, 
Mediterranean \textit{M. galloprovincialis}, \textit{M. trossulus}, \textit{M. californianus}.
Transcriptomes were translated using the \texttt{seqkit} program (v2.2.0; \cite{Shen2016})
We used genomes and associated annotations of \textit{M. coruscus} (GCA\_017311375.1; \textcite{Yang2021}),
Southern European \textit{M. edulis} (GCA\_905397895.1; \textcite{Corrochano-Fraile2021}),
Atlantic \textit{M. galloprovincialis} (GCA\_900618805.1; \textcite{Gerdol2020}),
and American \textit{M. edulis} (GCA\_019925275.1; annotation as personal communication of Tiago Hori, PEIMSO).
For GCA\_019925275.1, protein sequences where retrieve from the fasta and gff files using the python module gff3tool (v2.1.0).

We used \texttt{Orthofinder} (v2.5.2; \cite{Emms2015, Emms2019}) to find orthogroups and orthologue genes.
Species tree was inferred using the MSA method of \texttt{Orthofinder} and \texttt{IQTREE} tree inference (\texttt{-M msa -T iqtree}) \parencite{Emms2018, Nguyen2015}.

%\begin{itemize}
%	\item Supernova v2.1.1 \href{https://doi.org/10.1101/gr.214874.116}{cite}
%	\item Longranger v2.2.2
%	\item proc10xG \url{https://github.com/ucdavis-bioinformatics/proc10xG} commit 7afbfcf
%	\item nubeam-dedup \url{https://github.com/daihang16/nubeamdedup} commit 25dd385 \href{https://doi.org/10.1093/bioinformatics/btaa112}{cite}
%	\item KMC \url{https://github.com/tbenavi1/KMC} commit 1df71f6
%	\item GenomeScope 2.0 \url{https://github.com/tbenavi1/genomescope2.0} commit 5034ed4 \href{https://doi.org/10.1038/s41467-020-14998-3}{cite}
%	\item (? Smudgeplot commit 950d7b5)
%	\item Purge\_Dups \url{https://github.com/dfguan/purge_dups} commit e1934bb \href{https://doi.org/10.1093/bioinformatics/btaa025}{cite}
%	\item AGOUTI \url{https://github.com/svm-zhang/AGOUTI} commit a7e65d6 \href{https://doi.org/10.1186/s13742-016-0136-3}{cite}
%	\item augustus v3.3.3 (for AGOUTI analysis) (\protect\hyperlink{ref-Stanke2008}{\textbf{Stanke2008?}})
%	\item Blobtoolkit v2.4.0 \href{https://doi.org/10.1534/g3.119.400908}{cite}
%	\item KAT v2.4.2 \href{https://doi.org/10.1093/bioinformatics/btw663}{cite}
%	\item fastp v0.20.1 (\protect\hyperlink{ref-Chen2018a}{\textbf{Chen2018a?}})
%	\item samtools v1.12 (\protect\hyperlink{ref-Li2009}{\textbf{Li2009?}})
%	\item minimap2 v2.17 (\protect\hyperlink{ref-LiLi2017}{\textbf{LiLi2017?}})
%	\item bwa-mem2 v2.2.1 (\protect\hyperlink{ref-VasimuddinMd2019}{\textbf{VasimuddinMd2019?}})
%	\item bedtools v2.29.2 (\protect\hyperlink{ref-Quinlan2010}{\textbf{Quinlan2010?}})
%	\item seqkit v0.13.2 (\protect\hyperlink{ref-Shen2016}{\textbf{Shen2016?}})
%	\item rcorrector v1.0.4
%	\item trim-galore v0.6.6 \href{https://github.com/FelixKrueger/TrimGalore}{cite} (Cutadapt + FastQC wrapper)
%	\item busco v5.1.1 (\protect\hyperlink{ref-Seppey2019}{\textbf{Seppey2019?}})
%	\item assembly\_stats v0.1.4 \href{https://doi.org/10.5281/zenodo.3968774}{cite}
%	\item ragtag v1.1.1 (\protect\hyperlink{ref-Alonge2019}{\textbf{Alonge2019?}})
%	\item pilon v1.24 (\protect\hyperlink{ref-Walker2014}{\textbf{Walker2014?}})
%	\item TETools DFAM container v1.3.1
%	\item RepeatModeler v2.0.1 (\protect\hyperlink{ref-Flynn2020}{\textbf{Flynn2020?}})
%	\item cd-hit v4.8.1 (\protect\hyperlink{ref-Fu2012}{\textbf{Fu2012?}})
%	\item RepeatMasker v4.1.2-p1 (Smit, AFA, Hubley, R. RepeatModeler Open-1.0. 2008-2015 \url{http://www.repeatmasker.org}) (\protect\hyperlink{ref-Smit2013}{\textbf{Smit2013?}})
%\end{itemize}

\section{Results and discussion}

\subsection{Assembly statistics}

% Table 1
\begin{landscape}
	\include{figures/Tab1_asm_stats.tex}
\end{landscape}

\subsection{BUSCO scores}

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/Fig2_busco.pdf}
	\caption{Busco scores}
	\label{fig:busco}
\end{figure}

%\includegraphics{figures/full_figure_metazoa-1.pdf}

\section{Conclusion}\label{conclusion}

\section{Acknowledgement}

% References
\printbibliography

% Add supplementary here
\appendix

\end{document}

### 2022-04-07 scripts/01-asm_stats_prep.R

@@ -1,0 +1,52 @@
+library(tidyverse)
+library(xtable)
+
+asm_stats <- read_csv("data/assembly_stats.csv", col_types = cols())
+
+# name replacements
+asm_stats$asm <- str_replace_all(
+	asm_stats$asm, 
+	c("gallo_v"="MgalMED_v", 
+		"edu_v"="MeduEUS_v",
+		"tros_v"="MeduEUN_v",
+		"coruscus"="Mcor",
+		"gallo_GCA"="MgalATL_GCA",
+		"edu_GCA"="MeduEUS_GCA",
+		"eduam"="MeduAM")
+)
+
+write_tsv(asm_stats, "manuscript/supplementary_data/assembly_stats.tsv")
+
+asm_show <- c(
+	"Mcor_GCA017311375",
+	"MgalATL_GCA900618805",
+	"MeduEUS_GCA905397895",
+	"MeduAM_GCA019925415",
+	"MgalMED_v7",
+	"MeduEUS_v7",
+	"MeduEUN_v7"
+)
+
+asm_stats_table <- asm_stats %>% 
+	select(asm,
+				 C.L50, C.N50, C.median, C.sequence_count,
+				 S.L50, S.N50, S.gc_content, S.median, S.sequence_count, S.total_bps) %>% 
+	filter(asm %in% asm_show) %>% 
+	mutate(asm = fct_relevel(asm, asm_show)) %>% 
+	rename(assembly = asm)
+	
+
+tab_caption <- "
+Assembly statistics comparisons. C for contigs and S for scaffolds.
+"
+
+print(xtable(asm_stats_table,
+						 caption = tab_caption,
+						 label = "tab:asm_stats",
+						 digits = c(0,0,0,0,0,0,0,0,1,0,0,0)),
+	booktabs = TRUE,
+	caption.placement = "top",
+	include.rownames = FALSE,
+	scalebox = 0.7,
+	file = "manuscript/figures/Tab1_asm_stats.tex"
+)

### 2022-04-07 scripts/02-busco_figures.R

@@ -1,0 +1,79 @@
+library(tidyverse)
+library(patchwork)
+
+busco <- read_tsv("data/busco_summary.tsv", col_types = cols()) %>% 
+	mutate(asm = str_replace(asm, "_metazoa_odb10|_mollusca_odb10", ""))
+
+busco$asm <- str_replace_all(
+	busco$asm, 
+	c("gallo_v"="MgalMED_v", 
+		"edu_v"="MeduEUS_v",
+		"tros_v"="MeduEUN_v",
+		"coruscus"="Mcor",
+		"gallo_GCA"="MgalATL_GCA",
+		"edu_GCA"="MeduEUS_GCA",
+		"eduam"="MeduAM")
+)
+
+asm_show <- c(
+	"Mcor_GCA017311375",
+	"MgalATL_GCA900618805",
+	"MeduEUS_GCA905397895",
+	"MeduAM_GCA019925415",
+	"MgalMED_v7",
+	"MeduEUS_v7",
+	"MeduEUN_v7"
+)
+
+busco <- busco %>% # clean asm names
+	mutate(cat = fct_relevel(cat, "M", "F", "CD", "CS")) %>%
+	mutate(asm = fct_relevel(asm, asm_show) %>% 
+				 	fct_rev()) %>% 
+	group_by(asm, db) %>% 
+	mutate(proportion = N/cur_data()$N[cur_data()$cat=="T"]) %>% 
+	mutate(pct_fmt = paste(format(round(proportion*100, digits = 2)), "%"))
+
+
+fig_mollusca <- busco %>% 
+	filter(db == "mollusca_odb10") %>% 
+	filter(cat %in% c("CS", "CD", "F", "M")) %>%
+	filter(asm %in% asm_show) %>% 
+	ggplot(., aes(x = N, y = asm, fill = cat)) + 
+	geom_bar(position = "fill", stat = "identity", width = 0.9) +
+	scale_fill_manual(values = c("#D55E00", "#E69F00", "#56B4E9", "#009E73"),
+										guide = guide_legend(reverse=T),
+										name = "categories", 
+										labels = c("Missing", "Fragmented", "Complete duplicated", "Complete single copy")) +
+	scale_x_continuous(expand=c(0,0), labels = scales::percent, breaks = c(0,0.2,0.4,0.6,0.8,1)) + 
+	scale_y_discrete(expand=c(0,0)) +
+	geom_text(aes(label = N), position = position_fill(vjust = .5), size = 3) +
[…]

### 2022-04-07 scripts/03-species_tree.R

@@ -1,0 +1,9 @@
+library(treedataverse)
+
+tree <- read.newick("data/SpeciesTree_rooted_node_labels.txt")
+
+ggtree(tree) + 
+	geom_tiplab() +
+	geom_treescale(x=0, y=10) +
+	xlim(c(0,0.1)) +
+	theme_tree()

### 2022-05-17 manuscript/ms_genomes.tex
unknown changes

@@
 We used \texttt{Orthofinder} (v2.5.2; \cite{Emms2015, Emms2019}) to find orthogroups and orthologue genes.
-Species tree was inferred using the MSA method of \texttt{Orthofinder} and \texttt{IQTREE} tree inference (\texttt{-M msa -T iqtree}) \parencite{Emms2018, Nguyen2015}.
+Species tree was inferred using the MSA method of \texttt{Orthofinder} \parencite{Emms2018} and \texttt{IQTREE} (v2.2.0\_beta; \cite{Minh2020}) tree inference.
 

### 2022-06-17 manuscript/ms_genomes.tex
additions

@@
 
-Therefore, need for diverse references among the complex...
+As an effort to diversify the genomic resources available for the \textit{Mytilus} species complex,
+we assembled and annotate the genomes of three lineages using the 10X chromium technology.
+While our assemblies were initially fragmented due to the high level of heterozygosity,
+we leveraged the existence of a chromosome scale assembly of a sister species to scaffold our assemblies against it.
+We obtained assemblies equivalent to published ones in term of completeness for three new lineages of the species complex using a low sequencing budget and publicly available data.
+The resources produced and the assembly pipeline are freely available for use by the community.
 
@@
 
-We used the program \texttt{AGOUTI} (\url{https://github.com/svm-zhang/AGOUTI}; commit a7e65d6; \cite{Zhang2016a}) to improve scaffolding by using paired end RNA-seq reads.
+We used \texttt{AGOUTI} (\url{https://github.com/svm-zhang/AGOUTI}; commit a7e65d6; \cite{Zhang2016a}) to improve scaffolding by using paired end RNA-seq reads.
 For each species, a different set of published transcriptomes were used (Supplementary Table [[...]] for accessions).
@@
 Finally, gene completion analyses were carried out using \texttt{BUSCO} (v5.1.1; \cite{Manni2021}).
-We a Metazoan (\texttt{metazoa\_odb10.2021-02-24}) and a Molluscan database (\texttt{mollusca\_odb10.2020-08-05}) to assess and compare new and published assemblies.
-Published assemblies are:
+We used both a Metazoan (\texttt{metazoa\_odb10.2021-02-24}) and Molluscan (\texttt{mollusca\_odb10.2020-08-05}) database to assess and compare new and published assemblies.
+We compared our assemblies to the following published ones:
 \begin{itemize}
@@
 
-\subsection{Assembly statistics}
+The pre-assembly k-mer analysis carried out using Genomescope (using 21-mers) showed that, as expected, the genomes were highly heterozygous with values going from 3.49 to 8.77~\% (\cref{fig:genomescope}).
+Genomescope provide a bad fit and estimation in the case of the MeduEUS data due to a reduced coverage of the genome in that case, compared to the two other assemblies.
+
+Highly heterozygous genomes brings the risk of having a large number of duplicate contigs due to the separate assembly of the two alleles from a same locus.
+For this reason, we used the \texttt{purge\_dups} pipeline to try removing a maximum of such bias.
+This procedure reduced the number of complete duplicated genes in all assemblies (\cref{supfig:busco}, v1 to v4).
+
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{figures/Fig1_genomescope.pdf}
+	\caption{Genomescope}
+	\label{fig:genomescope}
+\end{figure}
 
@@
 
-\subsection{BUSCO scores}
+To assess the completeness of our assemblies, we compared them to four \textit{Mytilus} published assemblies on the basis of a Metazoan and a Molluscan set of single copy orthologous genes using BUSCO (\cref{fig:busco}).
+Overall, we show that our assemblies are equivalent to the published ones in terms of completeness.
 
@@
 
-\section{Conclusion}\label{conclusion}
+%\section{Conclusion}\label{conclusion}
 
-\section{Acknowledgement}
+\section*{Acknowledgements}
 
@@
 % Add supplementary here
+\newpage
 \appendix
+\setcounter{table}{0}
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\section*{Supplementary Information}
+
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{figures/supfig_busco.pdf}
+	\caption{Busco scores}
+	\label{supfig:busco}
+\end{figure}
 

### 2022-06-17 scripts/02-busco_figures.R

@@ -2,18 +2,24 @@ library(tidyverse)
 library(patchwork)
 
 busco <- read_tsv("data/busco_summary.tsv", col_types = cols()) %>% 
-	mutate(asm = str_replace(asm, "_metazoa_odb10|_mollusca_odb10", ""))
-
-busco$asm <- str_replace_all(
-	busco$asm, 
-	c("gallo_v"="MgalMED_v", 
-		"edu_v"="MeduEUS_v",
-		"tros_v"="MeduEUN_v",
-		"coruscus"="Mcor",
-		"gallo_GCA"="MgalATL_GCA",
-		"edu_GCA"="MeduEUS_GCA",
-		"eduam"="MeduAM")
-)
+	mutate(asm = str_replace(asm, "_metazoa_odb10|_mollusca_odb10", "")) %>% 
+	mutate(cat = fct_relevel(cat, "M", "F", "CD", "CS")) %>% 
+	group_by(asm, db) %>% mutate(perc = N/N[cat=="T"]*100) %>% 
+	ungroup %>% 
+	mutate(asm = str_replace_all(
+		asm, 
+		c("gallo_v"="MgalMED_v", 
+			"edu_v"="MeduEUS_v",
+			"tros_v"="MeduEUN_v",
+			"coruscus"="Mcor",
+			"gallo_GCA"="MgalATL_GCA",
+			"edu_GCA"="MeduEUS_GCA",
+			"eduam"="MeduAM")
+	)) %>% 
+	filter(!grepl("v2|v3", asm)) %>% 
+	mutate(asm = as_factor(asm),
+				 db = as_factor(db))
+	# mutate(label = paste0(N, '\n(', round(perc, digits = 1), ' %)'))
 
 asm_show <- c(
 	"Mcor_GCA017311375",
@@ -25,8 +31,7 @@ asm_show <- c(
 	"MeduEUN_v7"
 )
 
-busco <- busco %>% # clean asm names
-	mutate(cat = fct_relevel(cat, "M", "F", "CD", "CS")) %>%
+busco_fig2 <- busco %>% # clean asm names
 	mutate(asm = fct_relevel(asm, asm_show) %>% 
 				 	fct_rev()) %>% 
 	group_by(asm, db) %>% 
@@ -34,7 +39,7 @@ busco <- busco %>% # clean asm names
 	mutate(pct_fmt = paste(format(round(proportion*100, digits = 2)), "%"))
 
 
[…]

### 2022-06-17 manuscript/ms_genomes.tex
blobtoolkit legends

@@
 
-% Add supplementary here
+%=============================================
+% Supp
 \newpage
@@
 
+\subsection*{blob legends}
+
+None-binned blob plot of base coverage in GM against GC proportion for scaffolds in assembly MgalMED\_v7.
+Scaffolds are coloured by phylum and binned at a resolution of 30 divisions on each axis.
+Coloured nones within each bin are sized in proportion to the sum of individual scaffold lengths on a square-root scale, ranging from 1,005 to 1,127,105,406.
+Histograms show the distribution of scaffold length sum along each axis.
+
+None-binned blob plot of base coverage in GM against GC proportion for scaffolds in assembly MeduEUS\_v7.
+Scaffolds are coloured by phylum and binned at a resolution of 30 divisions on each axis.
+Coloured nones within each bin are sized in proportion to the sum of individual scaffold lengths on a square-root scale, ranging from 987 to 487,517,891.
+Histograms show the distribution of scaffold length sum along each axis.
+
+None-binned blob plot of base coverage in GM against GC proportion for scaffolds in assembly MeduEUN\_v7.
+Scaffolds are coloured by phylum and binned at a resolution of 30 divisions on each axis.
+Coloured nones within each bin are sized in proportion to the sum of individual scaffold lengths on a square-root scale, ranging from 1,011 to 1,136,301,196.
+Histograms show the distribution of scaffold length sum along each axis.
+
+\subsection*{cumulative legends}
+
+Cumulative scaffold length for assembly MgalMED\_v7.
+The grey line shows cumulative length for all scaffolds.
+Coloured lines show cumulative lengths of scaffolds assigned to each phylum using the bestsumorder taxrule.
+
+Cumulative scaffold length for assembly MeduEUS\_v7.
+The grey line shows cumulative length for all scaffolds.
+Coloured lines show cumulative lengths of scaffolds assigned to each phylum using the bestsumorder taxrule.
+
+Cumulative scaffold length for assembly MeduEUN\_v7.
+The grey line shows cumulative length for all scaffolds.
+Coloured lines show cumulative lengths of scaffolds assigned to each phylum using the bestsumorder taxrule.
+
+\subsection*{snailplot legends}
+
+Snail plot summary of assembly statistics for assembly MgalMED\_v7.
+The main plot is divided into 1,000 size-ordered bins around the circumference with each bin representing 0.1\% of the 1,658,656,017 bp assembly.
+The distribution of scaffold lengths is shown in dark grey with the plot radius scaled to the longest scaffold present in the assembly (104,729,878 bp, shown in red).
+Orange and pale-orange arcs show the N50 and N90 scaffold lengths (71,952,646 and 16,011 bp), respectively.
+The pale grey spiral shows the cumulative scaffold count on a log scale with white scale lines showing successive orders of magnitude.
+The blue and pale-blue area around the outside of the plot shows the distribution of GC, AT and N percentages in the same bins as the inner plot.
+A summary of complete, fragmented, duplicated and missing BUSCO genes in the mollusca\_odb10 set is shown in the top right.
+
+Snail plot summary of assembly statistics for assembly MeduEUS\_v7.
+The main plot is divided into 1,000 size-ordered bins around the circumference with each bin representing 0.1\% of the 2,076,685,641 bp assembly.
+The distribution of scaffold lengths is shown in dark grey with the plot radius scaled to the longest scaffold present in the assembly (88,305,666 bp, shown in red).
+Orange and pale-orange arcs show the N50 and N90 scaffold lengths (228,263 and 10,773 bp), respectively.
+The pale grey spiral shows the cumulative scaffold count on a log scale with white scale lines showing successive orders of magnitude.
+The blue and pale-blue area around the outside of the plot shows the distribution of GC, AT and N percentages in the same bins as the inner plot.
+A summary of complete, fragmented, duplicated and missing BUSCO genes in the mollusca\_odb10 set is shown in the top right.
+
+Snail plot summary of assembly statistics for assembly MeduEUN\_v7.
+The main plot is divided into 1,000 size-ordered bins around the circumference with each bin representing 0.1\% of the 1,764,246,486 bp assembly.
+The distribution of scaffold lengths is shown in dark grey with the plot radius scaled to the longest scaffold present in the assembly (110,194,685 bp, shown in red).
+Orange and pale-orange arcs show the N50 and N90 scaffold lengths (77,102,752 and 30,649 bp), respectively.
+The pale grey spiral shows the cumulative scaffold count on a log scale with white scale lines showing successive orders of magnitude.
+The blue and pale-blue area around the outside of the plot shows the distribution of GC, AT and N percentages in the same bins as the inner plot.
+A summary of complete, fragmented, duplicated and missing BUSCO genes in the mollusca\_odb10 set is shown in the top right.
+
 \end{document}

### 2022-06-17 manuscript/ms_genomes.tex
add btk figures

@@
 
-\subsection*{blob legends}
+\subsection*{Blobtoolkit}
 
-None-binned blob plot of base coverage in GM against GC proportion for scaffolds in assembly MgalMED\_v7.
+% blobs
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_blob_MgalMED_v7}
+	\caption{None-binned blob plot of base coverage in GM against GC proportion for scaffolds in assembly MgalMED\_v7.
 Scaffolds are coloured by phylum and binned at a resolution of 30 divisions on each axis.
 Coloured nones within each bin are sized in proportion to the sum of individual scaffold lengths on a square-root scale, ranging from 1,005 to 1,127,105,406.
-Histograms show the distribution of scaffold length sum along each axis.
+Histograms show the distribution of scaffold length sum along each axis.}
+	\label{supfig:btk-blob-MgalMED}
+\end{figure}
 
-None-binned blob plot of base coverage in GM against GC proportion for scaffolds in assembly MeduEUS\_v7.
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_blob_MgalMED_v7}
+	\caption{None-binned blob plot of base coverage in GM against GC proportion for scaffolds in assembly MeduEUS\_v7.
 Scaffolds are coloured by phylum and binned at a resolution of 30 divisions on each axis.
 Coloured nones within each bin are sized in proportion to the sum of individual scaffold lengths on a square-root scale, ranging from 987 to 487,517,891.
-Histograms show the distribution of scaffold length sum along each axis.
+Histograms show the distribution of scaffold length sum along each axis.}
+	\label{supfig:btk-blob-MeduEUS}
+\end{figure}
 
-None-binned blob plot of base coverage in GM against GC proportion for scaffolds in assembly MeduEUN\_v7.
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_blob_MgalMED_v7}
+	\caption{None-binned blob plot of base coverage in GM against GC proportion for scaffolds in assembly MeduEUN\_v7.
 Scaffolds are coloured by phylum and binned at a resolution of 30 divisions on each axis.
 Coloured nones within each bin are sized in proportion to the sum of individual scaffold lengths on a square-root scale, ranging from 1,011 to 1,136,301,196.
-Histograms show the distribution of scaffold length sum along each axis.
+Histograms show the distribution of scaffold length sum along each axis.}
+	\label{supfig:btk-blob-MeduEUN}
+\end{figure}
 
-\subsection*{cumulative legends}
+% cumulative lengths
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_cumulative_MgalMED_v7}
+	\caption{Cumulative scaffold length for assembly MgalMED\_v7.
+The grey line shows cumulative length for all scaffolds.
+Coloured lines show cumulative lengths of scaffolds assigned to each phylum using the bestsumorder taxrule.}
+	\label{supfig:btk-cumul-MgalMED}
+\end{figure}
 
-Cumulative scaffold length for assembly MgalMED\_v7.
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_cumulative_MeduEUS_v7}
+	\caption{Cumulative scaffold length for assembly MeduEUS\_v7.
 The grey line shows cumulative length for all scaffolds.
-Coloured lines show cumulative lengths of scaffolds assigned to each phylum using the bestsumorder taxrule.
+Coloured lines show cumulative lengths of scaffolds assigned to each phylum using the bestsumorder taxrule.}
+	\label{supfig:btk-cumul-MeduEUS}
+\end{figure}
 
-Cumulative scaffold length for assembly MeduEUS\_v7.
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_cumulative_MeduEUN_v7}
+	\caption{Cumulative scaffold length for assembly MeduEUN\_v7.
 The grey line shows cumulative length for all scaffolds.
-Coloured lines show cumulative lengths of scaffolds assigned to each phylum using the bestsumorder taxrule.
+Coloured lines show cumulative lengths of scaffolds assigned to each phylum using the bestsumorder taxrule.}
+	\label{supfig:btk-cumul-MeduEUN}
+\end{figure}
 
-Cumulative scaffold length for assembly MeduEUN\_v7.
-The grey line shows cumulative length for all scaffolds.
-Coloured lines show cumulative lengths of scaffolds assigned to each phylum using the bestsumorder taxrule.
-
-\subsection*{snailplot legends}
-
-Snail plot summary of assembly statistics for assembly MgalMED\_v7.
+% snail plots
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_snail_MgalMED_v7}
+	\caption{Snail plot summary of assembly statistics for assembly MgalMED\_v7.
 The main plot is divided into 1,000 size-ordered bins around the circumference with each bin representing 0.1\% of the 1,658,656,017 bp assembly.
@@
 The blue and pale-blue area around the outside of the plot shows the distribution of GC, AT and N percentages in the same bins as the inner plot.
-A summary of complete, fragmented, duplicated and missing BUSCO genes in the mollusca\_odb10 set is shown in the top right.
+A summary of complete, fragmented, duplicated and missing BUSCO genes in the mollusca\_odb10 set is shown in the top right.}
+	\label{supfig:btk-snail-MgalMED}
+\end{figure}
 
-Snail plot summary of assembly statistics for assembly MeduEUS\_v7.
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_snail_MeduEUS_v7}
+	\caption{Snail plot summary of assembly statistics for assembly MeduEUS\_v7.
 The main plot is divided into 1,000 size-ordered bins around the circumference with each bin representing 0.1\% of the 2,076,685,641 bp assembly.
@@
 A summary of complete, fragmented, duplicated and missing BUSCO genes in the mollusca\_odb10 set is shown in the top right.
+}
+	\label{supfig:btk-snail-MeduEUS}
+\end{figure}
 
-Snail plot summary of assembly statistics for assembly MeduEUN\_v7.
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_snail_MeduEUN_v7}
+	\caption{Snail plot summary of assembly statistics for assembly MeduEUN\_v7.
 The main plot is divided into 1,000 size-ordered bins around the circumference with each bin representing 0.1\% of the 1,764,246,486 bp assembly.
@@
 The blue and pale-blue area around the outside of the plot shows the distribution of GC, AT and N percentages in the same bins as the inner plot.
-A summary of complete, fragmented, duplicated and missing BUSCO genes in the mollusca\_odb10 set is shown in the top right.
+A summary of complete, fragmented, duplicated and missing BUSCO genes in the mollusca\_odb10 set is shown in the top right.}
+	\label{supfig:btk-snail-MeduEUN}
+\end{figure}
 

### 2022-06-20 manuscript/ms_genomes.tex
version table (mainly)

--- 2022-06-17 manuscript/ms_genomes.tex
@@
 
+\begin{table}[h]
+	\caption{Steps carried out for each version of assembly}
+	\label{suptab:versions}
+	\begin{tabular}{rll}
+		Version & Step & \\ \toprule
+		\rowcolor{gray!20}
+		v1 & Raw assembly from Supernova & \\
+		v2 & Supernova assembly using a filtered dataset of reads (\cref{preproc}) & \multirow{2}{*}{\rule{1pt}{25pt} Unused} \\
+		v3 & Assembly with purged duplicates (\texttt{purge\_dups}) from v3 & \\
+		\rowcolor{gray!20}
+		v4 & Assembly with purged duplicates (\texttt{purge\_dups}) from v1 & \\
+		\rowcolor{gray!20}
+		v5 & \texttt{Agouti} repaired scaffolds from v4 & \\
+		\rowcolor{gray!20}
+		v6 & Assembly with filtered contamination using Blobtoolkit results & \\
+		\rowcolor{gray!20}
+		v7 & Scaffolded assembly on \textit{M. coruscus} & \\ \bottomrule
+	\end{tabular}
+\end{table}
+\vspace{3em}
+
 \begin{figure}[h]
 	\includegraphics[width=\linewidth]{figures/supfig_busco.pdf}
-	\caption{Busco scores}
+	\caption{Busco scores on the metazoa\_odb10 (left panels) and mollusca\_odb10 (right panels) databases for each assembly MgalMED, MeduEUS and MeduEUN (top to bottom panels) across several versions (v1, v4, v5, v6, v7).}
 	\label{supfig:busco}
@@
 
+% blobs
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_blob_MgalMED_v7}
+	\caption{Blobtoolkit blob plot of base coverage in GM against GC proportion for scaffolds in assembly MgalMED\_v7.
+Scaffolds are colored by phylum and binned at a resolution of 30 divisions on each axis.
+Colored squares within each bin are sized in proportion to the sum of individual scaffold lengths on a square-root scale, ranging from 1,005 to 1,127,105,406.
+Histograms show the distribution of scaffold length sum along each axis.}
+	\label{supfig:btk-blob-MgalMED}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_blob_MeduEUS_v7}
+	\caption{Blobtoolkit blob plot of base coverage in GM against GC proportion for scaffolds in assembly MeduEUS\_v7.
+Scaffolds are colored by phylum and binned at a resolution of 30 divisions on each axis.
+Colored squares within each bin are sized in proportion to the sum of individual scaffold lengths on a square-root scale, ranging from 987 to 487,517,891.
+Histograms show the distribution of scaffold length sum along each axis.}
+	\label{supfig:btk-blob-MeduEUS}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_blob_MeduEUN_v7}
+	\caption{Blobtoolkit blob plot of base coverage in GM against GC proportion for scaffolds in assembly MeduEUN\_v7.
+Scaffolds are colored by phylum and binned at a resolution of 30 divisions on each axis.
+Colored squares within each bin are sized in proportion to the sum of individual scaffold lengths on a square-root scale, ranging from 1,011 to 1,136,301,196.
+Histograms show the distribution of scaffold length sum along each axis.}
+	\label{supfig:btk-blob-MeduEUN}
+\end{figure}
+
+% cumulative lengths
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_cumulative_MgalMED_v7}
+	\caption{Blobtoolkit cumulative scaffold length for assembly MgalMED\_v7.
+The gray line shows cumulative length for all scaffolds.
+Colored lines show cumulative lengths of scaffolds assigned to each phylum using the bestsumorder taxrule.}
+	\label{supfig:btk-cumul-MgalMED}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_cumulative_MeduEUS_v7}
+	\caption{Blobtoolkit cumulative scaffold length for assembly MeduEUS\_v7.
+The gray line shows cumulative length for all scaffolds.
+Colored lines show cumulative lengths of scaffolds assigned to each phylum using the bestsumorder taxrule.}
+	\label{supfig:btk-cumul-MeduEUS}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_cumulative_MeduEUN_v7}
+	\caption{Blobtoolkit cumulative scaffold length for assembly MeduEUN\_v7.
+The gray line shows cumulative length for all scaffolds.
+Colored lines show cumulative lengths of scaffolds assigned to each phylum using the bestsumorder taxrule.}
+	\label{supfig:btk-cumul-MeduEUN}
+\end{figure}
+
+% snail plots
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_snail_MgalMED_v7}
+	\caption{Blobtoolkit snail plot summary of assembly statistics for assembly MgalMED\_v7.
+The main plot is divided into 1,000 size-ordered bins around the circumference with each bin representing 0.1\% of the 1,658,656,017 bp assembly.
+The distribution of scaffold lengths is shown in dark gray with the plot radius scaled to the longest scaffold present in the assembly (104,729,878 bp, shown in red).
+Orange and pale-orange arcs show the N50 and N90 scaffold lengths (71,952,646 and 16,011 bp), respectively.
+The pale gray spiral shows the cumulative scaffold count on a log scale with white scale lines showing successive orders of magnitude.
+The blue and pale-blue area around the outside of the plot shows the distribution of GC, AT and N percentages in the same bins as the inner plot.
+A summary of complete, fragmented, duplicated and missing BUSCO genes in the mollusca\_odb10 set is shown in the top right.}
+	\label{supfig:btk-snail-MgalMED}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_snail_MeduEUS_v7}
+	\caption{Blobtoolkit snail plot summary of assembly statistics for assembly MeduEUS\_v7.
+The main plot is divided into 1,000 size-ordered bins around the circumference with each bin representing 0.1\% of the 2,076,685,641 bp assembly.
+The distribution of scaffold lengths is shown in dark gray with the plot radius scaled to the longest scaffold present in the assembly (88,305,666 bp, shown in red).
+Orange and pale-orange arcs show the N50 and N90 scaffold lengths (228,263 and 10,773 bp), respectively.
+The pale gray spiral shows the cumulative scaffold count on a log scale with white scale lines showing successive orders of magnitude.
+The blue and pale-blue area around the outside of the plot shows the distribution of GC, AT and N percentages in the same bins as the inner plot.
+A summary of complete, fragmented, duplicated and missing BUSCO genes in the mollusca\_odb10 set is shown in the top right.
+}
+	\label{supfig:btk-snail-MeduEUS}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/btk_snail_MeduEUN_v7}
+	\caption{Blobtoolkit snail plot summary of assembly statistics for assembly MeduEUN\_v7.
+The main plot is divided into 1,000 size-ordered bins around the circumference with each bin representing 0.1\% of the 1,764,246,486 bp assembly.
+The distribution of scaffold lengths is shown in dark gray with the plot radius scaled to the longest scaffold present in the assembly (110,194,685 bp, shown in red).
+Orange and pale-orange arcs show the N50 and N90 scaffold lengths (77,102,752 and 30,649 bp), respectively.
+The pale gray spiral shows the cumulative scaffold count on a log scale with white scale lines showing successive orders of magnitude.
+The blue and pale-blue area around the outside of the plot shows the distribution of GC, AT and N percentages in the same bins as the inner plot.
+A summary of complete, fragmented, duplicated and missing BUSCO genes in the mollusca\_odb10 set is shown in the top right.}
+	\label{supfig:btk-snail-MeduEUN}
+\end{figure}
+
 \end{document}

### 2022-06-20 manuscript/ms_genomes.tex
barcode filt

@@
 We then used \texttt{proc10xG} (\url{https://github.com/ucdavis-bioinformatics/proc10xG}; commit 7afbfcf) to split the reads from their 10X barcodes for further processing.
-We used a custom filtering step designed to remove under- and over-represented barcodes and their associated reads from the data (\texttt{filter\_barcodes.R} script, see \cite{Bary\&Gagnaire??}).
+We used a custom filtering step designed to remove under- and over-represented barcodes and their associated reads from the data (\cref{supfig:preproc_MgalMED,supfig:preproc_MeduEUS,supfig:preproc_MeduEUN}, \texttt{filter\_barcodes.R} script, see \cite{Bary\&Gagnaire??}).
 [[Add supfigs of filter\_barcodes]].
@@
 Preprocessed 10X reads (\cref{preproc}) were used to first estimate estimate genome size and heterozygosities of the three individuals.
-We used the reference free k-mer based method \texttt{genomescope} (\url{https://github.com/tbenavi1/genomescope2.0}; commit 5034ed4; \cite{Ranallo-Benavidez2020}) and the fork of the \texttt{KMC} k-mer  counting program (\url{https://github.com/tbenavi1/KMC}; commit 1df71f6).
+We used the reference free k-mer based method \texttt{GenomeScope} (\url{https://github.com/tbenavi1/genomescope2.0}; commit 5034ed4; \cite{Ranallo-Benavidez2020}) and the fork of the \texttt{KMC} k-mer  counting program (\url{https://github.com/tbenavi1/KMC}; commit 1df71f6).
 
@@
 
-The pre-assembly k-mer analysis carried out using Genomescope (using 21-mers) showed that, as expected, the genomes were highly heterozygous with values going from 3.49 to 8.77~\% (\cref{fig:genomescope}).
-Genomescope provide a bad fit and estimation in the case of the MeduEUS data due to a reduced coverage of the genome in that case, compared to the two other assemblies.
+The pre-assembly k-mer analysis carried out using GenomeScope (using 21-mers) showed that, as expected, the genomes were highly heterozygous with values going from 3.49 to 8.77~\% (\cref{fig:genomescope}).
+GenomeScope provide a bad fit and estimation in the case of the MeduEUS data due to a reduced coverage of the genome in that case, compared to the two other assemblies.
 
@@
 	\includegraphics[width=\linewidth]{figures/Fig1_genomescope.pdf}
-	\caption{Genomescope}
+	\caption{k-mer profile plots computed with 21-mers using GenomeScope.
+		Coverage histogram of the k-mers in blue.
+		Lines represent the fit of the GenomeScope models.
+		len: inferred genome length; uniq: percentage of the genome that is unique, aa: overall homozigosity; ab: overall heterozygosity; kcov: mean k-mer coverage for heterozygous bases; err: reads error rate; dup: average rate of read duplication; k: k-mer size; p: ploidy.
+	}
 	\label{fig:genomescope}
@@
 
+\subsection*{Tables}
+
 \begin{table}[h]
@@
 
+\subsection*{Figures}
+% preproc barcodes
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{figures/MgalMED_preproc_barcode_filt.pdf}
+	\caption{Histogram of read pairs identified for each 10X barcode for MgalMED.
+		As part of preprocessing step, barcodes for which too few or too many read pairs are associated with each unique barcode are removed from the dataset (red regions).}
+	\label{supfig:preproc_MgalMED}
+\end{figure}
+
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{figures/MeduEUS_preproc_barcode_filt.pdf}
+	\caption{Histogram of read pairs identified for each 10X barcode for MeduEUS.
+		As part of preprocessing step, barcodes for which too few or too many read pairs are associated with each unique barcode are removed from the dataset (red regions).}
+	\label{supfig:preproc_MeduEUS}
+\end{figure}
+
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{figures/MeduEUN_preproc_barcode_filt.pdf}
+	\caption{Histogram of read pairs identified for each 10X barcode for MeduEUN.
+	As part of preprocessing step, barcodes for which too few or too many read pairs are associated with each unique barcode are removed from the dataset (red regions).}
+	\label{supfig:preproc_MeduEUN}
+\end{figure}
+
 \begin{figure}[h]
 	\includegraphics[width=\linewidth]{figures/supfig_busco.pdf}
-	\caption{Busco scores on the metazoa\_odb10 (left panels) and mollusca\_odb10 (right panels) databases for each assembly MgalMED, MeduEUS and MeduEUN (top to bottom panels) across several versions (v1, v4, v5, v6, v7).}
+	\caption{Busco scores on the metazoa\_odb10 (left panels) and mollusca\_odb10 (right panels) databases for each assembly MgalMED, MeduEUS and MeduEUN (top to bottom panels) across several assembly versions (v1, v4, v5, v6, v7).}
 	\label{supfig:busco}

### 2022-06-22 manuscript/ms_genomes.tex
species tree

@@
 
+The species tree (\cref{fig:tree}) shows that the assemblies and the published genomes and transcriptomes are clustering as expected.
+\texttt{Orthofinder} \comment{provides}{provide the orthogroups somewhere} a resource of single-copy orthogroups across a large part the \textit{Mytilus} species complex.
+
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{figures/Fig3_tree.pdf}
+	\caption{Species tree using 1300 orthogroups with minimum of 81.8\% of species having single-copy genes in any orthogroup.
+	Color coding -- dark blue: \textit{M. edulis} Europe, light blue: \textit{M. edulis} America,
+	red: \textit{M. galloprovincialis} Mediterranean Sea, yellow: \textit{M. galloprovincialis} Atlantic,
+	green: \textit{M. trossulus}, pink: \textit{M. coruscus}, gray: \textit{M. californianus}.
+	}
+	\label{fig:tree}
+\end{figure}
+
 %\includegraphics{figures/full_figure_metazoa-1.pdf}

### 2022-06-22 scripts/03-species_tree.R

@@ -1,9 +1,45 @@
 library(treedataverse)
+library(forcats)
+library(ggtext)
 
 tree <- read.newick("data/SpeciesTree_rooted_node_labels.txt")
 
-ggtree(tree) + 
-	geom_tiplab() +
+tree$tip.label[5] = "bold(MgalMED)"
+tree$tip.label[6] = "bold(MeduEUN)"
+tree$tip.label[8] = "bold(MeduEUS)"
+
+lineages = as_factor(c(
+	'M. californianus',
+	'M. trossulus',
+	'M. galloprovincialis Atlantic',
+	'M. galloprovincialis Mediterranean Sea',
+	'M. galloprovincialis Mediterranean Sea',
+	'M. edulis Europe North',
+	'M. edulis Europe South',
+	'M. edulis Europe South',
+	'M. edulis America',
+	'M. edulis America',
+	'M. coruscus'
+))
+lineages_2 = fct_collapse(lineages, 'M. edulis Europe'=c('M. edulis Europe North', 'M. edulis Europe South'))
+color_pal = c("#999999", "#009E73", "#F0E442", "#D55E00",
+							"#0072B2", "#56B4E9", "#CC79A7")
+cols = factor(as.integer(lineages_2), labels = color_pal)
+
+
+
+p = ggtree(tree) + 
+	geom_tiplab(align=T, linesize=.5, offset=.002, parse=T) +
 	geom_treescale(x=0, y=10) +
 	xlim(c(0,0.1)) +
+	geom_tippoint(color=cols, size=4) +
 	theme_tree()
+
+# p + geom_text(aes(label=node), hjust=-.3)
+# 
+# p + geom_highlight(node=20, fill='blue') +
+# 	geom_highlight(node=21, fill='gold') +
+# 	geom_highlight(node=16, fill='red') +
+# 	geom_highlight(node=2, fill='green')
+
+ggsave("manuscript/figures/Fig3_tree.pdf", width = 8, height=4)

### 2022-06-24 manuscript/ms_genomes.tex
update

@@
 
+We introduce here newly assembled genomes from three lineages of the \textit{Mytilus} species complex.
+With a limited budget of $\sim$3000€ per genome, we managed to produce draft genomes of enough quality to be useful
+in applications such as population genomics and genetics.
+The method of 10X chromium pseudo-long reads combined with scaffolding using published data provided a quality comparable to published assemblies for \textit{Mytilus} species.
+
 The pre-assembly k-mer analysis carried out using GenomeScope (using 21-mers) showed that, as expected, the genomes were highly heterozygous with values going from 3.49 to 8.77~\% (\cref{fig:genomescope}).
@@
 \section*{Acknowledgements}
+
+We thank [XXX].
+AS was supported by the ANR grant [XXX] and by the National Institute of General Medical Sciences of the National Institutes of Health (grants NIH R01 GM108779 and R35 GM136290 to Dr Graham Coop).
 

### 2022-06-29 manuscript/ms_genomes.tex
kat spectra + data availability

@@
 High Molecular weight genomic DNA was extracted following \textcite{Nakayama1994}.
-We used a 15 mL Phase Lock Gel Heavy extraction with 3 steps of phenol/chloroform/isoamylalcool (25/24/1) Tris pH 8,1 followed by 2 chloroform extractions.
+We used a 15 mL Phase Lock Gel Heavy extraction with 3 steps of phenol / chloroform / isoamylalcool (25/24/1) Tris pH 8,1 followed by 2 chloroform extractions.
 After the last extraction, the aqueous supernatant was precipitated with 2 volumes of 100\% EtOH and the pellet was hooked from the solution with a sterile glass Pasteur pipette.
@@
 We then used \texttt{proc10xG} (\url{https://github.com/ucdavis-bioinformatics/proc10xG}; commit 7afbfcf) to split the reads from their 10X barcodes for further processing.
-We used a custom filtering step designed to remove under- and over-represented barcodes and their associated reads from the data (\cref{supfig:preproc_MgalMED,supfig:preproc_MeduEUS,supfig:preproc_MeduEUN}, \texttt{filter\_barcodes.R} script, see \cite{Bary\&Gagnaire??}).
-[[Add supfigs of filter\_barcodes]].
+We used a custom filtering step designed to remove under- and over-repre\-sented barcodes and their associated reads from the data (\cref{supfig:preproc_MgalMED,supfig:preproc_MeduEUS,supfig:preproc_MeduEUN}, \texttt{filter\_barcodes.R} script, see \cite{Bary\&Gagnaire??}).
 Reads were additionally filtered with \texttt{fastp} (v0.20.1; \cite{Chen2018}) with the main objective to remove poly-G tails created by the Illumina Novaseq sequencing technology.
@@
 We used \texttt{Augustus} (v3.3.3; \cite{Stanke2008}) to produce an intermediate annotation for each assembly using the \textit{Caenorhabditis} model.
-RNA-seq reads were first cleaned using \texttt{rcorrector} (v1.0.4; \cite{Song2015}) and \texttt{trimgalore} (v0.6.6; \url{https://www.bioinformatics.babraham.ac.uk/projects/trim_galore}; \texttt{--quality 20 --stringency 1 -e 0.1 --length 70}).
+RNA-seq reads were first cleaned using \texttt{rcorrector} (v1.0.4; \cite{Song2015}) and \texttt{trimgalore} (v0.6.6; \url{https://www.bioinformatics.babraham.ac.uk/projects/trim_galore}; \texttt{--quality 20 \\--stringency 1 -e 0.1 --length 70}).
 RNA-seq reads were mapped independently using \texttt{bwa-mem2} (v2.2.1; \cite{VasimuddinMd2019}), and then merged as a common bam file for each assembly with \texttt{samtools} (v1.12; \cite{Li2009}).
-Finally, the \texttt{AGOUTI} scaffolding pipeline was run using the gene prediction and mapped RNA-seq reads (-minMQ 20 -maxFracMM 0.05; using \texttt{python} v2.7.15 and \texttt{samtools} v1.10 for compatibility).
+Finally, the \texttt{AGOUTI} scaffolding pipeline was run using the gene prediction and mapped RNA-seq reads (\texttt{-minMQ 20 -maxFracMM 0.05}; using \texttt{python} v2.7.15 and \texttt{samtools} v1.10 for compatibility).
 
-As a last step, we ran \textit{Blobtoolkit} (v2.4.0; \cite{Challis2020}) on the three assemblies to evaluate quality and potential contamination levels.
-We used a custom script (\texttt{btk\_conta\_extraction.py}) to filter the assembly contigs based on the taxids found by the \textit{Blobtoolkit} pipeline to remove the most obvious contaminations.
+As a last step, we ran \texttt{Blobtoolkit} (v2.4.0; \cite{Challis2020}) on the three assemblies to evaluate quality and potential contamination levels.
+We used a custom script (\texttt{btk\_conta\_extraction.py}) to filter the assembly contigs based on the taxids found by the \texttt{Blobtoolkit} pipeline to remove the most obvious contaminations.
 Contigs matching taxids associated with viruses, bacteria and non-mollusca eukaryotes were removed.
@@
 
-We used \texttt{Orthofinder} (v2.5.2; \cite{Emms2015, Emms2019}) to find orthogroups and orthologue genes.
-Species tree was inferred using the MSA method of \texttt{Orthofinder} \parencite{Emms2018} and \texttt{IQTREE} (v2.2.0\_beta; \cite{Minh2020}) tree inference.
+We used \texttt{OrthoFinder} (v2.5.2; \cite{Emms2015, Emms2019}) to find orthogroups and orthologue genes.
+Species tree was inferred using the MSA method of \texttt{OrthoFinder} \parencite{Emms2018} and \texttt{IQTREE} (v2.2.0\_beta; \cite{Minh2020}) tree inference.
 
@@
 This procedure reduced the number of complete duplicated genes in all assemblies (\cref{supfig:busco}, v1 to v4).
+Overall, the compared \texttt{KAT} spectra analyses show that the assembly steps reduced the amount of duplication in all assemblies (\cref{supfig:kat_spectra}).
 
@@
 The species tree (\cref{fig:tree}) shows that the assemblies and the published genomes and transcriptomes are clustering as expected.
-\texttt{Orthofinder} \comment{provides}{provide the orthogroups somewhere} a resource of single-copy orthogroups across a large part the \textit{Mytilus} species complex.
+\texttt{OrthoFinder} \comment{provides}{provide the orthogroups somewhere} a resource of single-copy orthogroups across a large part the \textit{Mytilus} species complex.
+
+OrthoFinder results:
+OrthoFinder assigned 499004 genes (91.9\% of total) to 50543 orthogroups. Fifty percent of all 
+genes were in orthogroups with 13 or more genes (G50 was 13) and were contained in the largest
+11982 orthogroups (O50 was 11982). There were 1852 orthogroups with all species present and 
+173 of these consisted entirely of single-copy genes.
+
+Species tree using 1300 orthogroups with minimum of 81.8\% of species having single-copy genes in any orthogroup (value automatically selected by \texttt{OrthoFinder}).
 
@@
 
-%\includegraphics{figures/full_figure_metazoa-1.pdf}
+%\section{Conclusion}\label{conclusion}
 
-%\section{Conclusion}\label{conclusion}
+\section*{Data availability}
+
+Raw reads are available under BioProject PRJNA785550.
+Assemblies are available under accessions JAKGDF000000000 (MgalMED), JAKGDG000000000 (MeduEUS), and JAKGDH000000000 (MeduEUN).
+Annotations are available under the Zenodo archive [XXX].
+The assembly pipeline is available at \url{https://github.com/alxsimon/assembly_10x}.
+The OrthoFinder pipeline is available at [XXX].
 
@@
 \begin{figure}[h]
-	\includegraphics[width=\linewidth]{figures/MgalMED_preproc_barcode_filt.pdf}
+	\includegraphics[width=\linewidth]{figures/supfig_MgalMED_preproc_barcode_filt.pdf}
 	\caption{Histogram of read pairs identified for each 10X barcode for MgalMED.
@@
 \begin{figure}[h]
-	\includegraphics[width=\linewidth]{figures/MeduEUS_preproc_barcode_filt.pdf}
+	\includegraphics[width=\linewidth]{figures/supfig_MeduEUS_preproc_barcode_filt.pdf}
 	\caption{Histogram of read pairs identified for each 10X barcode for MeduEUS.
@@
 \begin{figure}[h]
-	\includegraphics[width=\linewidth]{figures/MeduEUN_preproc_barcode_filt.pdf}
+	\includegraphics[width=\linewidth]{figures/supfig_MeduEUN_preproc_barcode_filt.pdf}
 	\caption{Histogram of read pairs identified for each 10X barcode for MeduEUN.
@@
 	\label{supfig:preproc_MeduEUN}
+\end{figure}
+
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{figures/supfig_kat_spectra.pdf}
+	\caption{KAT comparison of k-mer spectra between assembly and preprocessed reads.
+	Results are compared between the initial assembly (v1, left column) and the final assembly (v7, right column) for each sample MgalMED, MeduEUS and MeduEUN (from top to bottom).}
+	\label{supfig:kat_spectra}
 \end{figure}

### 2022-06-30 manuscript/ms_genomes.tex
repeats (mainly)

@@
 We used \texttt{AGOUTI} (\url{https://github.com/svm-zhang/AGOUTI}; commit a7e65d6; \cite{Zhang2016a}) to improve scaffolding by using paired end RNA-seq reads.
-For each species, a different set of published transcriptomes were used (Supplementary Table [[...]] for accessions).
+For each species, a different set of published transcriptomes were used (See Supplementary File 1 for accession numbers).
 \texttt{AGOUTI} require a gene prediction as input in addition to RNA-seq reads.
@@
 
-To obtain an annotation for each assembly, we used both database and RNA-seq information.
-We used \texttt{Braker2} (v2.1.6; \cite{Bruna2021}) as an initial step, using both a protein database and RNA-seq reads (preprocessed in \cref{draft_asm}).
+We used \texttt{Braker2} (v2.1.6; \cite{Bruna2021}) as to obtain structural annotations, using both a protein database and RNA-seq reads (preprocessed in \cref{draft_asm}).
 To build the protein database, we used all \textit{Mollusca} proteins (taxid 6447) from OrthoDB (v10.1; \cite{Kriventseva2019}).
@@
 
-We used the \texttt{Mantis} pipeline (\url{https://github.com/PedroMTQ/mantis}; commit c6cb597; \cite{Queiros2021}) to obtain consensus annotations of genes based on multiple databases.
+We used the \texttt{Mantis} pipeline (\url{https://github.com/PedroMTQ/mantis}; commit c6cb597; \cite{Queiros2021}) to obtain consensus functional annotations of genes based on multiple databases.
 Protein sequences for each assembly were built from the \texttt{Braker2} annotation using the \texttt{python} module \texttt{gff3tool} (v2).
@@
 
-We used \texttt{OrthoFinder} (v2.5.2; \cite{Emms2015, Emms2019}) to find orthogroups and orthologue genes.
-Species tree was inferred using the MSA method of \texttt{OrthoFinder} \parencite{Emms2018} and \texttt{IQTREE} (v2.2.0\_beta; \cite{Minh2020}) tree inference.
+We used \texttt{OrthoFinder} (v2.5.4; \cite{Emms2015, Emms2019}) to find orthogroups and orthologue genes.
+Species tree was inferred using the MSA method of \texttt{OrthoFinder} \parencite{Emms2018} using the \texttt{MAFFT} aligner (v7.505; \cite{Katoh2013}), the \texttt{STRIDE} species tree rooting algorithm \parencite{Emms2017} and the \texttt{FastTree} software for tree inference (v2.1.11; \cite{Price2009}).
 
@@
 
+\subsection{Assembly results}
+
 We introduce here newly assembled genomes from three lineages of the \textit{Mytilus} species complex.
-With a limited budget of $\sim$3000€ per genome, we managed to produce draft genomes of enough quality to be useful
+With a limited budget of $\sim3000$€ per genome, we managed to produce draft genomes of enough quality to be useful
 in applications such as population genomics and genetics.
@@
 
-The pre-assembly k-mer analysis carried out using GenomeScope (using 21-mers) showed that, as expected, the genomes were highly heterozygous with values going from 3.49 to 8.77~\% (\cref{fig:genomescope}).
-GenomeScope provide a bad fit and estimation in the case of the MeduEUS data due to a reduced coverage of the genome in that case, compared to the two other assemblies.
+The pre-assembly k-mer analysis carried out using GenomeScope (using 21-mers) showed that, as expected, the genomes were highly heterozygous with values going from $3.49$ to $8.77$~\% (\cref{fig:genomescope}).
+For the MeduEUS data, GenomeScope provides a fit of bad quality and an estimation probably off due to a reduced slightly lower sequencing depth compared to the two other assemblies.
+In that case, the heterozygous peak was not identifiable.
 
@@
 
+Stats:
+GC contents: 32.1\%, 34.6\%, 32.1\%
+
 % Table 1
@@
 
-The species tree (\cref{fig:tree}) shows that the assemblies and the published genomes and transcriptomes are clustering as expected.
-\texttt{OrthoFinder} \comment{provides}{provide the orthogroups somewhere} a resource of single-copy orthogroups across a large part the \textit{Mytilus} species complex.
+The MeduEUS assembly suffered from an increased level of contamination compared to the other two assemblies.
+Despite a broad contaminant filtering step, the Busco blob analysis (\cref{supfig:btk-blob-MeduEUS})
+shows a second GC content peak centered around 40\% (above the ~32\% peaks of this and other genomes).
 
-OrthoFinder results:
+\subsection{Repeat contents}
+
+More than half of each genome was identified as repeats and masked by \texttt{RepeatMasker}.
+Repeats are estimated to amount to 57.22, 61.10 and 55.53\% of the bases of MgalMED, MeduEUS and MeduEUN assemblies respectively.
+The majority of repeats are unclassified, followed by retroelements with a balanced contribution of LINEs and LTR elements
+(detailed \texttt{RepeatMasker} results in \cref{suptab:repeats}).
+
+\subsection{Annotations}
+
+[[Here goes annotation stats]]
+
+\subsection{Species tree}
+
 OrthoFinder assigned 499004 genes (91.9\% of total) to 50543 orthogroups. Fifty percent of all 
@@
 
-Species tree using 1300 orthogroups with minimum of 81.8\% of species having single-copy genes in any orthogroup (value automatically selected by \texttt{OrthoFinder}).
+The species tree was built using 1300 orthogroups with a minimum of 81.8\% of species having single-copy genes in any orthogroup (\cref{fig:tree}).
+It shows that the assemblies and the published genomes and transcriptomes are clustering as expected.
 
@@
 	\includegraphics[width=\linewidth]{figures/Fig3_tree.pdf}
-	\caption{Species tree using 1300 orthogroups with minimum of 81.8\% of species having single-copy genes in any orthogroup.
+	\caption{Species tree using 1300 orthogroups with a minimum of 81.8\% of species having single-copy genes in any orthogroup.
 	Color coding -- dark blue: \textit{M. edulis} Europe, light blue: \textit{M. edulis} America,
@@
 
-Raw reads are available under BioProject PRJNA785550.
+Raw data are available under BioProject PRJNA785550.
 Assemblies are available under accessions JAKGDF000000000 (MgalMED), JAKGDG000000000 (MeduEUS), and JAKGDH000000000 (MeduEUN).
-Annotations are available under the Zenodo archive [XXX].
 The assembly pipeline is available at \url{https://github.com/alxsimon/assembly_10x}.
-The OrthoFinder pipeline is available at [XXX].
+Annotations and the OrthoFinder pipeline and results are available on the Zenodo archive [[XXX]].
 
@@
 
-We thank [XXX].
-AS was supported by the ANR grant [XXX] and by the National Institute of General Medical Sciences of the National Institutes of Health (grants NIH R01 GM108779 and R35 GM136290 to Dr Graham Coop).
+We thank [[XXX]].
+We thank IFB core cluster for computational resources [[...]]
+We thank Tiago Horri for providing his annotation of the American \textit{Mytilus edulis} genome (GCA\_019925275.1).
+AS was supported by the ANR grant [[XXX]] and by the National Institute of General Medical Sciences of the National Institutes of Health (grants NIH R01 GM108779 and R35 GM136290 to Dr Graham Coop).
 
@@
 
+% Table contains discrepancies of the sums but this is what RepeatMasker outputs... So I don't know...
+\include{figures/Suptab_repeats.tex}
+
 \end{document}

### 2022-06-30 scripts/01-asm_stats_prep.R

@@ -43,10 +43,11 @@ Assembly statistics comparisons. C for contigs and S for scaffolds.
 print(xtable(asm_stats_table,
 						 caption = tab_caption,
 						 label = "tab:asm_stats",
-						 digits = c(0,0,0,0,0,0,0,0,1,0,0,0)),
+						 digits = c(0,0,-2,-2,-2,0,0,-2,1,-2,0,-2)),
 	booktabs = TRUE,
 	caption.placement = "top",
 	include.rownames = FALSE,
-	scalebox = 0.7,
+	scalebox = 0.8,
+	math.style.exponents = TRUE,
 	file = "manuscript/figures/Tab1_asm_stats.tex"
 )

### 2022-07-08 manuscript/ms_genomes.tex
include mantis db and change version

@@
 
-We used the \texttt{Mantis} pipeline (\url{https://github.com/PedroMTQ/mantis}; commit c6cb597; \cite{Queiros2021}) to obtain consensus functional annotations of genes based on multiple databases.
-Protein sequences for each assembly were built from the \texttt{Braker2} annotation using the \texttt{python} module \texttt{gff3tool} (v2).
+We used the \texttt{Mantis} pipeline (v1.5.5; \cite{Queiros2021}) to obtain consensus functional annotations of genes based on multiple databases.
+Protein sequences for each assembly were built from the structural annotations using the \texttt{python} module \texttt{gff3tool} (v2).
+\texttt{Mantis} was run with default parameters and databases:
+kofam \parencite{Aramaki2020}, NPFM \parencite{Lu2020}, eggNOG \parencite{Huerta-Cepas2019}, pfam \parencite{El-Gebali2019}, and tcdb \parencite{Saier2021}.
 
@@
 Species tree was inferred using the MSA method of \texttt{OrthoFinder} \parencite{Emms2018} using the \texttt{MAFFT} aligner (v7.505; \cite{Katoh2013}), the \texttt{STRIDE} species tree rooting algorithm \parencite{Emms2017} and the \texttt{FastTree} software for tree inference (v2.1.11; \cite{Price2009}).
-
-%\begin{itemize}
-%	\item Supernova v2.1.1 \href{https://doi.org/10.1101/gr.214874.116}{cite}
-%	\item Longranger v2.2.2
-%	\item proc10xG \url{https://github.com/ucdavis-bioinformatics/proc10xG} commit 7afbfcf
-%	\item nubeam-dedup \url{https://github.com/daihang16/nubeamdedup} commit 25dd385 \href{https://doi.org/10.1093/bioinformatics/btaa112}{cite}
-%	\item KMC \url{https://github.com/tbenavi1/KMC} commit 1df71f6
-%	\item GenomeScope 2.0 \url{https://github.com/tbenavi1/genomescope2.0} commit 5034ed4 \href{https://doi.org/10.1038/s41467-020-14998-3}{cite}
-%	\item (? Smudgeplot commit 950d7b5)
-%	\item Purge\_Dups \url{https://github.com/dfguan/purge_dups} commit e1934bb \href{https://doi.org/10.1093/bioinformatics/btaa025}{cite}
-%	\item AGOUTI \url{https://github.com/svm-zhang/AGOUTI} commit a7e65d6 \href{https://doi.org/10.1186/s13742-016-0136-3}{cite}
-%	\item augustus v3.3.3 (for AGOUTI analysis) (\protect\hyperlink{ref-Stanke2008}{\textbf{Stanke2008?}})
-%	\item Blobtoolkit v2.4.0 \href{https://doi.org/10.1534/g3.119.400908}{cite}
-%	\item KAT v2.4.2 \href{https://doi.org/10.1093/bioinformatics/btw663}{cite}
-%	\item fastp v0.20.1 (\protect\hyperlink{ref-Chen2018a}{\textbf{Chen2018a?}})
-%	\item samtools v1.12 (\protect\hyperlink{ref-Li2009}{\textbf{Li2009?}})
-%	\item minimap2 v2.17 (\protect\hyperlink{ref-LiLi2017}{\textbf{LiLi2017?}})
-%	\item bwa-mem2 v2.2.1 (\protect\hyperlink{ref-VasimuddinMd2019}{\textbf{VasimuddinMd2019?}})
-%	\item bedtools v2.29.2 (\protect\hyperlink{ref-Quinlan2010}{\textbf{Quinlan2010?}})
-%	\item seqkit v0.13.2 (\protect\hyperlink{ref-Shen2016}{\textbf{Shen2016?}})
-%	\item rcorrector v1.0.4
-%	\item trim-galore v0.6.6 \href{https://github.com/FelixKrueger/TrimGalore}{cite} (Cutadapt + FastQC wrapper)
-%	\item busco v5.1.1 (\protect\hyperlink{ref-Seppey2019}{\textbf{Seppey2019?}})
-%	\item assembly\_stats v0.1.4 \href{https://doi.org/10.5281/zenodo.3968774}{cite}
-%	\item ragtag v1.1.1 (\protect\hyperlink{ref-Alonge2019}{\textbf{Alonge2019?}})
-%	\item pilon v1.24 (\protect\hyperlink{ref-Walker2014}{\textbf{Walker2014?}})
-%	\item TETools DFAM container v1.3.1
-%	\item RepeatModeler v2.0.1 (\protect\hyperlink{ref-Flynn2020}{\textbf{Flynn2020?}})
-%	\item cd-hit v4.8.1 (\protect\hyperlink{ref-Fu2012}{\textbf{Fu2012?}})
-%	\item RepeatMasker v4.1.2-p1 (Smit, AFA, Hubley, R. RepeatModeler Open-1.0. 2008-2015 \url{http://www.repeatmasker.org}) (\protect\hyperlink{ref-Smit2013}{\textbf{Smit2013?}})
-%\end{itemize}
 
@@
 
-\subsection{Annotations}
-
-[[Here goes annotation stats]]
+%\subsection{Annotations}
+%
+%[[Here goes annotation stats]]
 

### 2022-08-24 manuscript/ms_genomes.tex
Nico

@@
 
-\title{Draft genomes of the three northern hemisphere blue mussel lineages: North and South European \emph{Mytilus edulis} and Mediterranean \emph{Mytilus galloprovincialis}.}
+\title{Draft genomes of three northern hemisphere blue mussel lineages: North and South European \emph{Mytilus edulis} and Mediterranean \emph{Mytilus galloprovincialis}.}
 
-\author[1,*]{Alexis Simon}
-\author[1]{Christine Arbiol}
-\author[1]{Nicolas Bierne}
+\author[1,2,*]{Alexis Simon}
 
-\affil[1]{CNRS, Univ Montpellier, \ldots{}}
-
-%
+\affil[1]{ISEM, EPHE, IRD, Université Montpellier, Montpellier, France}
+\affil[2]{Center of Population Biology and Department of Evolution and Ecology, University of California Davis, Davis, California, USA}
 
@@
 \begin{abstract}
-	Using the 10X chromium long reads technology, we provide draft genomes for three closely related blue mussel species from the \emph{Mytilus} species complex.
-	The objective was to produce affordable genomic resources for population and evolutionary genomic studies.
-	Genomes are fragmented but represent a large portion of the genome, with good sizes and BUSCO scores.
+    The blue mussel species complex (\textit{Mytilus edulis}) is of particular interest both as model species in population genetics and ecology,
+    but also as a economic resource in many regions.
+    We add here three new draft genomes to the fast increasing genomic resources of this complex.
+	Using 10X genomics reads, we assembled genomes of closely related blue mussel lineages from the \emph{Mytilus} species complex in the Northern hemisphere.
+	Given the huge diversity within and between lineages in this complex, 
+	the objective was to produce affordable genomic resources for population and evolutionary genomic studies to complement already published genomes.
+	%Genomes are fragmented but represent a large portion of the genome, with good sizes and BUSCO scores.
 \end{abstract}
@@
 \textit{Mytilus edulis}, \textit{M. galloprovincialis} and \textit{M. trossulus}.
-Within each species, evolutionary relevant lineages can be identified.
-Two lineages of \textit{M.
-galloprovincialis} can be identified, one Atlantic lineage and one Mediterranean lineage separated by a hybrid zone along the Almeria-Oran front \parencite{Quesada1995, Fraisse2016a, ElAyari2019}.
-Three lineages of \textit{M. edulis} can be identified: (i) an American lineage, (ii) a Southern European lineage, and (iii) a Northern European lineage \parencite{Fraisse2016a, Simon2020}.
+Within each species, evolutionary relevant lineages can be identified:
+two lineages in \textit{M.
+galloprovincialis}, (i)  an Atlantic lineage (MgalATL) and (ii) a Mediterranean lineage (MgalMED), separated by a hybrid zone at the Almeria-Oran front and Algerian coast \parencite{Quesada1995, Fraisse2016a, ElAyari2019}, and
+three lineages in \textit{M. edulis}, (i) an American lineage (MeduAM), (ii) a Southern European lineage (MeduEUS), and (iii) a Northern European lineage (MeduEUN) \parencite{Fraisse2016a, Simon2020}.
+
+Increased effort has been put in the last few years to produce genomic resources in the \textit{Mytilus} species complex.
+[[XXX]]
 
@@
 we leveraged the existence of a chromosome scale assembly of a sister species to scaffold our assemblies against it.
-We obtained assemblies equivalent to published ones in term of completeness for three new lineages of the species complex using a low sequencing budget and publicly available data.
+We obtained assemblies equivalent to published ones in term of completeness for three new lineages of the Mytilus species complex using a low sequencing budget and publicly available data.
 The resources produced and the assembly pipeline are freely available for use by the community.
@@
 \begin{itemize}
-	\item \emph{M. galloprovincialis} Mediterranean MgalMED; Sète, France
-	\item \emph{M. edulis} Southern European MeduEUS; ???
-	\item \emph{M. edulis} Northern Europe MeduEUN; ???
+	\item \emph{M. galloprovincialis} Mediterranean (MgalMED); Sète ($43^\circ24'35.3''N$, $3^\circ41'41.1''E$, Occitanie, France, Mediterranean Sea)
+	\item \emph{M. edulis} Southern European (MeduEUS); Agon-Coutainville ($49^\circ0'44.784''N$,\\$1^\circ35'55.643''W$, Normandy, France, English Channel)
+	\item \emph{M. edulis} Northern Europe (MeduEUN); Mishukovo ($69^\circ02'34.3''N$, $33^\circ01'45.9''E$, Kola Bay, Russia, Barents Sea)
 \end{itemize}
@@
 
-After this period of pre-treatment at ambient temperature, proteinase K was added at a final concentration of 150 \(\mu\)g/mL and the solution was incubated overnight at 56°C.
+After this period of pre-treatment at ambient temperature, proteinase K was added at a final concentration of 150 \(\mu\)g/mL and the solution was incubated overnight at $56^\circ$C.
 
 High Molecular weight genomic DNA was extracted following \textcite{Nakayama1994}.
-We used a 15 mL Phase Lock Gel Heavy extraction with 3 steps of phenol / chloroform / isoamylalcool (25/24/1) Tris pH 8,1 followed by 2 chloroform extractions.
+We used a 15 mL Phase Lock Gel Heavy extraction with three steps of extraction using phenol/chloroform/isoamyl alcohol (25/24/1 proportions) equilibrated with Tris-HCl to pH 8.1, followed by two chloroform extractions.
 After the last extraction, the aqueous supernatant was precipitated with 2 volumes of 100\% EtOH and the pellet was hooked from the solution with a sterile glass Pasteur pipette.
-The pellet was rinsed several times in 80\% EtOH before being dried at room temperature.
+The pellet was rinsed several times with 80\% EtOH before being dried at room temperature.
 
@@
 We then used \texttt{proc10xG} (\url{https://github.com/ucdavis-bioinformatics/proc10xG}; commit 7afbfcf) to split the reads from their 10X barcodes for further processing.
-We used a custom filtering step designed to remove under- and over-repre\-sented barcodes and their associated reads from the data (\cref{supfig:preproc_MgalMED,supfig:preproc_MeduEUS,supfig:preproc_MeduEUN}, \texttt{filter\_barcodes.R} script, see \cite{Bary\&Gagnaire??}).
+We used a custom filtering step designed to remove under- and over-repre\-sented barcodes and their associated reads from the data (\cref{supfig:preproc_MgalMED,supfig:preproc_MeduEUS,supfig:preproc_MeduEUN}, \texttt{filter\_barcodes.R} script, personal communication of P-A Gagnaire).
 Reads were additionally filtered with \texttt{fastp} (v0.20.1; \cite{Chen2018}) with the main objective to remove poly-G tails created by the Illumina Novaseq sequencing technology.
@@
 Contigs matching taxids associated with viruses, bacteria and non-mollusca eukaryotes were removed.
-More specifically contigs for were removed for virus contamination if they presented a hit percentage of more than 10~\% of their length.
+More specifically contigs were removed for virus contamination if they presented a hit percentage of more than 10~\% of their length.
 Contigs were removed for eukaryote contamination only when presenting only hits to taxa outside Mollusca on more than 10~\% of their length.
@@
 
-The NCBI submission process identified a few errors that needed correcting.
+The NCBI submission process identified a few errors that needed correction.
 A small number of adaptor sequences and duplicates were removed to comply with NCBI requirements
@@
 
-The pre-assembly k-mer analysis carried out using GenomeScope (using 21-mers) showed that, as expected, the genomes were highly heterozygous with values going from $3.49$ to $8.77$~\% (\cref{fig:genomescope}).
-For the MeduEUS data, GenomeScope provides a fit of bad quality and an estimation probably off due to a reduced slightly lower sequencing depth compared to the two other assemblies.
-In that case, the heterozygous peak was not identifiable.
+The pre-assembly k-mer analysis carried out using GenomeScope (using 21-mers) showed that, as expected, the genomes of MgalMED and MeduEUN were highly heterozygous with values of $3.49$ and $4.03$~\% respectively (\cref{fig:genomescope}).
+The heterozygous peak of MeduEUS was not identifiable due to lower sequencing depth and GenomeScope provides a bad quality fit that we did not interpret.
 
@@
 
-We thank [[XXX]].
-We thank IFB core cluster for computational resources [[...]]
+I thank Nicolas Bierne and Pierre-Alexandre Gagnaire for their help on the project and discussions.
+I thank Christine Arbiol for DNA extractions.
+Some bioinformatic analyses were performed on the Core Cluster of the Institut Français de Bioinformatique (ANR-11-INBS-0013).
 We thank Tiago Horri for providing his annotation of the American \textit{Mytilus edulis} genome (GCA\_019925275.1).
-AS was supported by the ANR grant [[XXX]] and by the National Institute of General Medical Sciences of the National Institutes of Health (grants NIH R01 GM108779 and R35 GM136290 to Dr Graham Coop).
+AS was supported by the French ANR grant TRANSCAN (ANR-18-CE35-0009) and by the National Institute of General Medical Sciences of the National Institutes of Health (grants NIH R01 GM108779 and R35 GM136290 to Graham Coop).
 

### 2022-08-30 manuscript/ms_genomes.tex
improve intro, abstract, add zenodo

@@
     The blue mussel species complex (\textit{Mytilus edulis}) is of particular interest both as model species in population genetics and ecology,
-    but also as a economic resource in many regions.
-    We add here three new draft genomes to the fast increasing genomic resources of this complex.
-	Using 10X genomics reads, we assembled genomes of closely related blue mussel lineages from the \emph{Mytilus} species complex in the Northern hemisphere.
+    but also as an economic resource in many regions.
+	Using 10X genomics pseudo long reads, I assembled genomes of closely related blue mussel lineages from the \emph{Mytilus} species complex in the Northern hemisphere.
 	Given the huge diversity within and between lineages in this complex, 
-	the objective was to produce affordable genomic resources for population and evolutionary genomic studies to complement already published genomes.
-	%Genomes are fragmented but represent a large portion of the genome, with good sizes and BUSCO scores.
+	the objective was to produce affordable genomic resources for population and evolutionary genomic studies to broaden the coverage of this diverse species complex.
+	I used transcriptome guided corrections and scaffolding on a chromosome scale genome of a close species to reduce the fragmentation of the genomes.
+	The result is a set of partially fragmented genomes of equivalent completeness to already published genomes.
+	Three new draft genomes are added to the fast increasing genomic resources of this complex for the Mediterranean \textit{M.
+galloprovincialis}, the South-European \textit{M.
+edulis} and the the North-European \textit{M.
+edulis}.
 \end{abstract}
@@
 The \textit{Mytilus} species complex has been a model system in population genetics, adaptation, hybridization and speciation since genetic variants could be identified \parencite{Milkman1970, Koehn1972, Ahmad1977, Skibinski1978, Quesada1995b, Bierne2003, Fraisse2016a, Simon2021}.
+This genus is also an important food source worldwide.
+While \textit{Mytilus} species are actively cultured,
+human selection has been limited in blue mussels as most of the production rely on naturally captured spat.
+Therefore, given the extreme diversity observed within and between species in this genus,
+extensive genomic resources are needed to both advance our fundamental knowledge and potentially accrue our capacity to
+improve its culture in the face of large environmental changes.
 
 The \textit{Mytilus} species complex is composed of three taxonomically recognized and partially reproductively isolated species in the northern hemisphere,
-\textit{Mytilus edulis}, \textit{M. galloprovincialis} and \textit{M. trossulus}.
+\textit{Mytilus edulis}, \textit{M.
+galloprovincialis} and \textit{M.
+trossulus}; and several other species in the southern hemisphere (\textit{M.
+chilensis}, \textit{M.
+platensis}, \textit{M.
+planulatus}).
+Between those species, hybridization is possible in contact zones, either natural or anthropogenic.
 Within each species, evolutionary relevant lineages can be identified:
@@
 three lineages in \textit{M. edulis}, (i) an American lineage (MeduAM), (ii) a Southern European lineage (MeduEUS), and (iii) a Northern European lineage (MeduEUN) \parencite{Fraisse2016a, Simon2020}.
+Two close species in the genus are of interest: \textit{M.
+coruscus}, a species found in Asia, and \textit{M.
+californianus}, a species found on the Pacific coast of North America.
+While they do not appear to readily hybridize with species of the \textit{Mytilus} species complex, they are of similar interest to the study of marine mussels.
 
 Increased effort has been put in the last few years to produce genomic resources in the \textit{Mytilus} species complex.
-[[XXX]]
+Quality genomes are available for the following lineages of \textit{Mytilus}: 
+Atlantic \textit{M. galloprovincialis} \parencite{Gerdol2020},
+Southern European \textit{M. edulis} \parencite{Corrochano-Fraile2022}, 
+\textit{M. coruscus} \parencite{Yang2021, Li2020}, 
+\textit{M. californianus} \parencite{Paggeot2022},
+and an American \textit{M. edulis} assembly (unpublished, accession GCA\_019925275).
+This increased effort has highlighted characteristics of interest in those species such as a high proportion of gene presence-absence \parencite{Gerdol2020} or evidence of genome duplication \parencite{Corrochano-Fraile2022}.
 
 As an effort to diversify the genomic resources available for the \textit{Mytilus} species complex,
-we assembled and annotate the genomes of three lineages using the 10X chromium technology.
-While our assemblies were initially fragmented due to the high level of heterozygosity,
-we leveraged the existence of a chromosome scale assembly of a sister species to scaffold our assemblies against it.
-We obtained assemblies equivalent to published ones in term of completeness for three new lineages of the Mytilus species complex using a low sequencing budget and publicly available data.
+I assembled and annotated the genomes of three lineages using the 10X chromium technology.
+While these assemblies were initially fragmented due to the high level of heterozygosity,
+I leveraged the existence of a chromosome scale assembly of a sister species to scaffold against it.
+I obtained assemblies equivalent to published ones in term of completeness for three new lineages of the Mytilus species complex using a low sequencing budget and publicly available data.
 The resources produced and the assembly pipeline are freely available for use by the community.
@@
 Therefore some assembly and annotation steps wrongly used \textit{M. trossulus} transcriptomes.
-While this is not ideal, we think results have not been strongly impacted by this issue.
+While this is not ideal, I think results have not been strongly impacted by this issue.
 
@@
 High Molecular weight genomic DNA was extracted following \textcite{Nakayama1994}.
-We used a 15 mL Phase Lock Gel Heavy extraction with three steps of extraction using phenol/chloroform/isoamyl alcohol (25/24/1 proportions) equilibrated with Tris-HCl to pH 8.1, followed by two chloroform extractions.
+A 15 mL Phase Lock Gel Heavy extraction was used with three steps of extraction using phenol/chloroform/isoamyl alcohol (25/24/1 proportions) equilibrated with Tris-HCl to pH 8.1, followed by two chloroform extractions.
 After the last extraction, the aqueous supernatant was precipitated with 2 volumes of 100\% EtOH and the pellet was hooked from the solution with a sterile glass Pasteur pipette.
@@
 
-We preprocessed 10X reads using the following pipeline for use in several algorithms (\cref{draft_asm}).
-We first removed duplicate reads using \texttt{Nubeam-dedup} (\cite{Dai2020}; commit 25dd385).
-We then used \texttt{proc10xG} (\url{https://github.com/ucdavis-bioinformatics/proc10xG}; commit 7afbfcf) to split the reads from their 10X barcodes for further processing.
-We used a custom filtering step designed to remove under- and over-repre\-sented barcodes and their associated reads from the data (\cref{supfig:preproc_MgalMED,supfig:preproc_MeduEUS,supfig:preproc_MeduEUN}, \texttt{filter\_barcodes.R} script, personal communication of P-A Gagnaire).
+I preprocessed 10X reads using the following pipeline for use in several algorithms (\cref{draft_asm}).
+I first removed duplicate reads using \texttt{Nubeam-dedup} (\cite{Dai2020}; commit 25dd385).
+I then used \texttt{proc10xG} (\url{https://github.com/ucdavis-bioinformatics/proc10xG}; commit 7afbfcf) to split the reads from their 10X barcodes for further processing.
+I used a custom filtering step designed to remove under- and over-repre\-sented barcodes and their associated reads from the data (\cref{supfig:preproc_MgalMED,supfig:preproc_MeduEUS,supfig:preproc_MeduEUN}, \texttt{filter\_barcodes.R} script, personal communication of P-A Gagnaire).
 Reads were additionally filtered with \texttt{fastp} (v0.20.1; \cite{Chen2018}) with the main objective to remove poly-G tails created by the Illumina Novaseq sequencing technology.
-We obtained at this point reads equivalent to a short-read sequencing run, usable in some parts of the assembly and quality control pipeline.
-Additionally, to obtain reads compatible to 10X genomics tools, we filtered and reassembled reads with their 10X barcodes in \texttt{proc10xG} using the \texttt{filter\_10xReads.py} and \texttt{regen\_10xReads.py} scripts.
+I obtained at this point reads equivalent to a short-read sequencing run, usable in some parts of the assembly and quality control pipeline.
+Additionally, to obtain reads compatible to 10X genomics tools, I filtered and reassembled reads with their 10X barcodes in \texttt{proc10xG} using the \texttt{filter\_10xReads.py} and \texttt{regen\_10xReads.py} scripts.
 
@@
 
-For each genome we used \texttt{Supernova} v2.1.1 \parencite{Weisenfeld2017} to assemble raw 10X reads.
-To avoid hard stops in \texttt{Supernova} due to both data quantity slightly under 10X genomics recommendations and an overestimation of genome size by Supernova, we used all available reads (\texttt{--maxreads='all'}) and accepted extreme coverage (\texttt{--accept-extreme-coverage}).
-We produced every style of Fasta output available in \texttt{Supernova} but only used the pseudo-haploid output in the following pipeline.
+For each genome I used \texttt{Supernova} v2.1.1 \parencite{Weisenfeld2017} to assemble raw 10X reads.
+To avoid hard stops in \texttt{Supernova} due to both data quantity slightly under 10X genomics recommendations and an overestimation of genome size by Supernova, I used all available reads (\texttt{--maxreads='all'}) and accepted extreme coverage (\texttt{--accept-extreme-coverage}).
+I produced every style of Fasta output available in \texttt{Supernova} but only used the pseudo-haploid output in the following pipeline.
 
-To remove duplicate haplotypes in the assemblies, we followed the \texttt{purge\_dups} pipeline (\cite{Guan2020}; commit e1934bb).
-We first used the \texttt{longranger} v2.2.2 align algorithm to map preprocessed reads (\cref{preproc}) to the pseudo-haploid genomes to use in the \texttt{ngscstat} step.
+To remove duplicate haplotypes in the assemblies, I followed the \texttt{purge\_dups} pipeline (\cite{Guan2020}; commit e1934bb).
+I first used the \texttt{longranger} v2.2.2 align algorithm to map preprocessed reads (\cref{preproc}) to the pseudo-haploid genomes to use in the \texttt{ngscstat} step.
 \texttt{Minimap2} (v2.17; \cite{Li2017}) was used in the contig self-mapping step.
-We obtained the purged assemblies using the \texttt{get\_seqs} steps without restricting the purging to the end of contigs (without the \texttt{-e} option).
+I obtained the purged assemblies using the \texttt{get\_seqs} steps without restricting the purging to the end of contigs (without the \texttt{-e} option).
 
-We used \texttt{AGOUTI} (\url{https://github.com/svm-zhang/AGOUTI}; commit a7e65d6; \cite{Zhang2016a}) to improve scaffolding by using paired end RNA-seq reads.
-For each species, a different set of published transcriptomes were used (See Supplementary File 1 for accession numbers).
+I used \texttt{AGOUTI} (\url{https://github.com/svm-zhang/AGOUTI}; commit a7e65d6; \cite{Zhang2016a}) to improve scaffolding by using paired end RNA-seq reads.
+For each species, a different set of published transcriptomes were used (see Supplementary File 1 for accession numbers).
 \texttt{AGOUTI} require a gene prediction as input in addition to RNA-seq reads.
-We used \texttt{Augustus} (v3.3.3; \cite{Stanke2008}) to produce an intermediate annotation for each assembly using the \textit{Caenorhabditis} model.
+I used \texttt{Augustus} (v3.3.3; \cite{Stanke2008}) to produce an intermediate annotation for each assembly using the \textit{Caenorhabditis} model.
 RNA-seq reads were first cleaned using \texttt{rcorrector} (v1.0.4; \cite{Song2015}) and \texttt{trimgalore} (v0.6.6; \url{https://www.bioinformatics.babraham.ac.uk/projects/trim_galore}; \texttt{--quality 20 \\--stringency 1 -e 0.1 --length 70}).
@@
 
-As a last step, we ran \texttt{Blobtoolkit} (v2.4.0; \cite{Challis2020}) on the three assemblies to evaluate quality and potential contamination levels.
-We used a custom script (\texttt{btk\_conta\_extraction.py}) to filter the assembly contigs based on the taxids found by the \texttt{Blobtoolkit} pipeline to remove the most obvious contaminations.
+As a last step, I ran \texttt{Blobtoolkit} (v2.4.0; \cite{Challis2020}) on the three assemblies to evaluate quality and potential contamination levels.
+I used a custom script (\texttt{btk\_conta\_extraction.py}) to filter the assembly contigs based on the taxids found by the \texttt{Blobtoolkit} pipeline to remove the most obvious contaminations.
 Contigs matching taxids associated with viruses, bacteria and non-mollusca eukaryotes were removed.
@@
 At the time of assembly, the closest relative of the \textit{Mytilus} species of interest with a chromosome scale assembly was \textit{Mytilus coruscus} (GCA\_017311375.1; \cite{Yang2021}).
-Given the conserved number of 14 chromosomes in the \textit{Mytilus} genus, we decided to scaffold our contigs on this high quality reference.
-Additionally, we had Oxford Nanopore reads for the MeduEUN individual. 
-We used \texttt{minimap2} (v2.17; \cite{Li2017}) and \texttt{LRScaf} (v1.1.10; \url{https://github.com/shingocat/lrscaf}) to first improve the MeduEUN assembly with this small amount of long reads.
+Given the conserved number of 14 chromosomes in the \textit{Mytilus} genus, I decided to scaffold our contigs on this high quality reference.
+Additionally, I had Oxford Nanopore reads for the MeduEUN individual. 
+I used \texttt{minimap2} (v2.17; \cite{Li2017}) and \texttt{LRScaf} (v1.1.10; \url{https://github.com/shingocat/lrscaf}) to first improve the MeduEUN assembly with this small amount of long reads.
 
-Then for each assembly, we ran the scaffolder \texttt{RagTag} (v1.1.1; \cite{Alonge2021}) to position contigs on the \emph{M. coruscus} chromosomal assembly.
+Then for each assembly, I ran the scaffolder \texttt{RagTag} (v1.1.1; \cite{Alonge2021}) to position contigs on the \emph{M. coruscus} chromosomal assembly.
 
@@
 \texttt{Pilon} attempts to improve the assembly based on mapped read information (gap filling and error corrections).
-We used \texttt{bwa-mem2} to map the debarcoded and filtered 10X reads (\cref{preproc}).
+I used \texttt{bwa-mem2} to map the debarcoded and filtered 10X reads (\cref{preproc}).
 In addition, Oxford Nanopore reads for MeduEUN were also used in \texttt{Pilon} for the given assembly.
@@
 
-We masked repeats for the purpose of annotation using \texttt{RepeatModeler} (v2.0.1; \cite{Flynn2020}) and \texttt{RepeatMasker} (v4.1.2-p1; \cite{Smit2013}) through the TETools DFAM container (v1.3.1; \url{https://hub.docker.com/r/dfam/tetools}).
-We first built repeat databases for each of five assemblies MgalMED, MeduEUS, MeduEUN, \textit{M.
+I masked repeats for the purpose of annotation using \texttt{RepeatModeler} (v2.0.1; \cite{Flynn2020}) and \texttt{RepeatMasker} (v4.1.2-p1; \cite{Smit2013}) through the TETools DFAM container (v1.3.1; \url{https://hub.docker.com/r/dfam/tetools}).
+I first built repeat databases for each of five assemblies MgalMED, MeduEUS, MeduEUN, \textit{M.
 coruscus} (GCA\_017311375.1; \cite{Yang2021}) and \textit{M.
 galloprovincialis} from the Atlantic (GCA\_900618805.1; \cite{Gerdol2020}).
-Then, we built a common \textit{Mytilus} database of repeats by merging those five databases using \texttt{cd-hit} (v4.8.1; \cite{Fu2012}).
-We used the same options as used by default in \texttt{RepeatModeler}: \texttt{-aS 0.8 -c 0.8 -g 1 -G 0 -A 80 -M 10000}.
+Then, I built a common \textit{Mytilus} database of repeats by merging those five databases using \texttt{cd-hit} (v4.8.1; \cite{Fu2012}).
+I used the same options as used by default in \texttt{RepeatModeler}: \texttt{-aS 0.8 -c 0.8 -g 1 -G 0 -A 80 -M 10000}.
 Finally, soft repeat masking was performed on the assemblies with \texttt{RepeatMasker} using the merged database.
@@
 
-We used \texttt{Braker2} (v2.1.6; \cite{Bruna2021}) as to obtain structural annotations, using both a protein database and RNA-seq reads (preprocessed in \cref{draft_asm}).
-To build the protein database, we used all \textit{Mollusca} proteins (taxid 6447) from OrthoDB (v10.1; \cite{Kriventseva2019}).
-To provide gene presence hints, we mapped all RNA-seq reads for each species using \texttt{HISAT2} (v2.2.1; \cite{Kim2019}).
+I used \texttt{Braker2} (v2.1.6; \cite{Bruna2021}) as to obtain structural annotations, using both a protein database and RNA-seq reads (preprocessed in \cref{draft_asm}).
+To build the protein database, I used all \textit{Mollusca} proteins (taxid 6447) from OrthoDB (v10.1; \cite{Kriventseva2019}).
+To provide gene presence hints, I mapped all RNA-seq reads for each species using \texttt{HISAT2} (v2.2.1; \cite{Kim2019}).
 
-We used the \texttt{Mantis} pipeline (v1.5.5; \cite{Queiros2021}) to obtain consensus functional annotations of genes based on multiple databases.
+I used the \texttt{Mantis} pipeline (v1.5.5; \cite{Queiros2021}) to obtain consensus functional annotations of genes based on multiple databases.
 Protein sequences for each assembly were built from the structural annotations using the \texttt{python} module \texttt{gff3tool} (v2).
@@
 Preprocessed 10X reads (\cref{preproc}) were used to first estimate estimate genome size and heterozygosities of the three individuals.
-We used the reference free k-mer based method \texttt{GenomeScope} (\url{https://github.com/tbenavi1/genomescope2.0}; commit 5034ed4; \cite{Ranallo-Benavidez2020}) and the fork of the \texttt{KMC} k-mer  counting program (\url{https://github.com/tbenavi1/KMC}; commit 1df71f6).
+I used the reference free k-mer based method \texttt{GenomeScope} (\url{https://github.com/tbenavi1/genomescope2.0}; commit 5034ed4; \cite{Ranallo-Benavidez2020}) and the fork of the \texttt{KMC} k-mer  counting program (\url{https://github.com/tbenavi1/KMC}; commit 1df71f6).
 
@@
 
-To assess the remaining level of duplication in the assemblies, we used the program \texttt{KAT} (v2.4.2; \cite{Mapleson2017}) to compare k-mer spectra from reads (preprocessed 10X) and from the assembly.
+To assess the remaining level of duplication in the assemblies, I used the program \texttt{KAT} (v2.4.2; \cite{Mapleson2017}) to compare k-mer spectra from reads (preprocessed 10X) and from the assembly.
 
 Finally, gene completion analyses were carried out using \texttt{BUSCO} (v5.1.1; \cite{Manni2021}).
-We used both a Metazoan (\texttt{metazoa\_odb10.2021-02-24}) and Molluscan (\texttt{mollusca\_odb10.2020-08-05}) database to assess and compare new and published assemblies.
-We compared our assemblies to the following published ones:
+I used both a Metazoan (\texttt{metazoa\_odb10.2021-02-24}) and Molluscan (\texttt{mollusca\_odb10.2020-08-05}) database to assess and compare new and published assemblies.
+I compared our assemblies to the following published ones:
 \begin{itemize}
@@
 	\item \textit{M. galloprovincialis} from the Altantic lineage, GCA\_900618805.1, \textcite{Gerdol2020};
-	\item \textit{M. edulis} from the Southern European lineage, GCA\_905397895.1, \textcite{Corrochano-Fraile2021}
+	\item \textit{M. edulis} from the Southern European lineage, GCA\_905397895.1, \textcite{Corrochano-Fraile2022}
 	\item \textit{M. edulis} from the American lineage, GCA\_019925275.1.
@@
 
-We compiled protein sequences from published \textit{Mytilus} genomes and transcriptomes,
+I compiled protein sequences from published \textit{Mytilus} genomes and transcriptomes,
 in addition to the current three genomes and annotations, to build a species tree.
-We used transcriptomes produced in \textcite{Popovic2020a} for American \textit{M. edulis}, 
+I used transcriptomes produced in \textcite{Popovic2020a} for American \textit{M. edulis}, 
 Mediterranean \textit{M. galloprovincialis}, \textit{M. trossulus}, \textit{M. californianus}.
 Transcriptomes were translated using the \texttt{seqkit} program (v2.2.0; \cite{Shen2016})
-We used genomes and associated annotations of \textit{M. coruscus} (GCA\_017311375.1; \textcite{Yang2021}),
-Southern European \textit{M. edulis} (GCA\_905397895.1; \textcite{Corrochano-Fraile2021}),
-Atlantic \textit{M. galloprovincialis} (GCA\_900618805.1; \textcite{Gerdol2020}),
+I used genomes and associated annotations of \textit{M. coruscus} (GCA\_017311375.1; \cite{Yang2021}),
+Southern European \textit{M. edulis} (GCA\_905397895.1; \cite{Corrochano-Fraile2022}),
+Atlantic \textit{M. galloprovincialis} (GCA\_900618805.1; \cite{Gerdol2020}),
 and American \textit{M. edulis} (GCA\_019925275.1; annotation as personal communication of Tiago Hori, PEIMSO).
@@
 
-We used \texttt{OrthoFinder} (v2.5.4; \cite{Emms2015, Emms2019}) to find orthogroups and orthologue genes.
+I used \texttt{OrthoFinder} (v2.5.4; \cite{Emms2015, Emms2019}) to find orthogroups and orthologue genes.
 Species tree was inferred using the MSA method of \texttt{OrthoFinder} \parencite{Emms2018} using the \texttt{MAFFT} aligner (v7.505; \cite{Katoh2013}), the \texttt{STRIDE} species tree rooting algorithm \parencite{Emms2017} and the \texttt{FastTree} software for tree inference (v2.1.11; \cite{Price2009}).
@@
 
-We introduce here newly assembled genomes from three lineages of the \textit{Mytilus} species complex.
-With a limited budget of $\sim3000$€ per genome, we managed to produce draft genomes of enough quality to be useful
+I introduce here newly assembled genomes from three lineages of the \textit{Mytilus} species complex.
+With a limited sequencing budget of $\sim3000$€ per genome, I managed to produce draft genomes of enough quality to be useful
 in applications such as population genomics and genetics.
@@
 The pre-assembly k-mer analysis carried out using GenomeScope (using 21-mers) showed that, as expected, the genomes of MgalMED and MeduEUN were highly heterozygous with values of $3.49$ and $4.03$~\% respectively (\cref{fig:genomescope}).
-The heterozygous peak of MeduEUS was not identifiable due to lower sequencing depth and GenomeScope provides a bad quality fit that we did not interpret.
+The heterozygous peak of MeduEUS was not identifiable due to lower sequencing depth and GenomeScope provides a bad quality fit that I did not interpret.
 
-Highly heterozygous genomes brings the risk of having a large number of duplicate contigs due to the separate assembly of the two alleles from a same locus.
-For this reason, we used the \texttt{purge\_dups} pipeline to try removing a maximum of such bias.
+Highly heterozygous genomes brings the risk of having a large number of duplicate contigs due to the separate assembly of the multiple alleles from given locus.
+For this reason, I used the \texttt{purge\_dups} pipeline to try removing a maximum of such bias.
 This procedure reduced the number of complete duplicated genes in all assemblies (\cref{supfig:busco}, v1 to v4).
@@
 
-Stats:
-GC contents: 32.1\%, 34.6\%, 32.1\%
+Assembly statistics for the three new assemblies and four publicly available assemblies are presented in \cref{tab:asm_stats} (complete statistics are presented in Supplementary File 2).
+With the exception of MeduEUS, GC content and genome sizes are close to what is found in other lineages.
+While the number of scaffolds for the new assemblies is still high, most of the genomes are contained in 14 large scaffolds corresponding to the number of chromosomes found in \textit{Mytilus} species.
+Residual scaffolds are sequences that could not be placed on the \textit{M. coruscus} genome,
+maybe due to inter-species presence-absence or too high divergence.
 
@@
 
-To assess the completeness of our assemblies, we compared them to four \textit{Mytilus} published assemblies on the basis of a Metazoan and a Molluscan set of single copy orthologous genes using BUSCO (\cref{fig:busco}).
-Overall, we show that our assemblies are equivalent to the published ones in terms of completeness.
+To assess the completeness of our assemblies, I compared them to four \textit{Mytilus} published assemblies on the basis of a Metazoan and a Molluscan set of single copy orthologous genes using BUSCO (\cref{fig:busco}).
+Overall, I show that our assemblies are equivalent to the published ones in terms of completeness.
 
@@
 shows a second GC content peak centered around 40\% (above the ~32\% peaks of this and other genomes).
+This contamination could for instance be due to bacteria present in the mussel if its physiological state was degraded before fixation.
+The metagenomic composition of contaminants was not investigated.
 
@@
 More than half of each genome was identified as repeats and masked by \texttt{RepeatMasker}.
-Repeats are estimated to amount to 57.22, 61.10 and 55.53\% of the bases of MgalMED, MeduEUS and MeduEUN assemblies respectively.
+Repeats are estimated to amount to 57.22, 61.10 and 55.53\% of MgalMED, MeduEUS and MeduEUN assemblies respectively.
 The majority of repeats are unclassified, followed by retroelements with a balanced contribution of LINEs and LTR elements
@@
 
-%\subsection{Annotations}
-%
-%[[Here goes annotation stats]]
+\subsection{Annotations}
+
+The assemblies were annotated using a consensus pipeline to provide additional resources for future uses.
+These annotation were used to produce the following species tree.
 
@@
 
-%\section{Conclusion}\label{conclusion}
-
 \section*{Data availability}
@@
 The assembly pipeline is available at \url{https://github.com/alxsimon/assembly_10x}.
-Annotations and the OrthoFinder pipeline and results are available on the Zenodo archive [[XXX]].
+Annotations, and the OrthoFinder pipeline and results are available in the Zenodo archive \href{https://doi.org/10.5281/zenodo.7034399}{10.5281/zenodo.7034399}.
+Data and code for this manuscript are available at \url{https://github.com/alxsimon/Mytilus_ref_genomes_paper}.
 
@@
 
-I thank Nicolas Bierne and Pierre-Alexandre Gagnaire for their help on the project and discussions.
-I thank Christine Arbiol for DNA extractions.
+I thank Nicolas Bierne and Pierre-Alexandre Gagnaire for their help on the project and discussions;
+Christine Arbiol for DNA extractions;
+Maurine Hammel and Erika Burioli for sharing their transcriptome data;
+and Petr Strelkov for sampling of the \textit{M. trossulus}.
 Some bioinformatic analyses were performed on the Core Cluster of the Institut Français de Bioinformatique (ANR-11-INBS-0013).
-We thank Tiago Horri for providing his annotation of the American \textit{Mytilus edulis} genome (GCA\_019925275.1).
+I thank Tiago Horri for providing his annotation of the American \textit{Mytilus edulis} genome (GCA\_019925275.1).
 AS was supported by the French ANR grant TRANSCAN (ANR-18-CE35-0009) and by the National Institute of General Medical Sciences of the National Institutes of Health (grants NIH R01 GM108779 and R35 GM136290 to Graham Coop).

### 2022-09-02 manuscript/ms_genomes.tex
change title

@@
 
-\title{Draft genomes of three northern hemisphere blue mussel lineages: North and South European \emph{Mytilus edulis} and Mediterranean \emph{Mytilus galloprovincialis}.}
+\title{Three new genome assemblies of blue mussel lineages: North and South European \emph{Mytilus edulis} and Mediterranean \emph{Mytilus galloprovincialis}.}
 
 \author[1,2,*]{Alexis Simon}
+\date{}
 
@@
     but also as an economic resource in many regions.
-	Using 10X genomics pseudo long reads, I assembled genomes of closely related blue mussel lineages from the \emph{Mytilus} species complex in the Northern hemisphere.
+	Using 10X genomics pseudo-long reads, I assembled genomes of three closely related blue mussel lineages from the \emph{Mytilus} species complex in the Northern hemisphere.
 	Given the huge diversity within and between lineages in this complex, 
@@
 	green: \textit{M. trossulus}, pink: \textit{M. coruscus}, gray: \textit{M. californianus}.
+	The \enquote{popovic} suffix indicate transcriptome date from \textcite{Popovic2020a}.
 	}
